---
title: "Direct ROC Comparison"
author: "Peter Fiorica"
date: "2025-10-20"
output: html_document
---

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(PRSmix))
suppressPackageStartupMessages(library(argparse))
suppressPackageStartupMessages(library(purrr))

"%&%"=function(a,b) paste(a,b,sep="")
```


# Introduction
These notes serve to respond to the third round of revisions for the JACC CardioOncology submission regarding PGS performance in breast cancer survivors.

>A main concern is the use of the Friedman test for Figure 4. The reviewers and our editors agree that more standard analyses should be used to analyze the data in Figure 4. The 95% CI of the AUC and values themselves do appear largely overlapping and the Board would like clarification on the use of the statistical test. In further consultation with our Statistical Editor and Reviewer 2, we would like to suggest a more standard statistical approach be used to test the added value when evaluating AUCs: https://doi.org/10.2307/2531595. 

# Analyses
```{r}
categorical_vars <- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd","prev_anycvd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

outcomes <- c("inc_anycmd", "inc_hyperten", "inc_diabetes", "inc_dyslipid", "inc_anycvd", "inc_seriouscvd", "inc_arrhythmia", "inc_hf", "inc_ischemic", "inc_stroke", "inc_vte", "inc_cvddeath")
cat_def <- c("M","M", "M","M", "V","V","V","V", "V", "V","V", "V")
dict <-setNames(cat_def, outcomes)

IID_pheno = "IID"
```

## Ischemic Heart Disease

```{r}
original_beta_files_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/janise_share/weights.txt"
isbinary <- TRUE
allPGS_list = NULL
metascore = NULL 
liabilityR2 = TRUE
train_size_list = NULL
training_result_file = NULL
read_pred_training = FALSE
read_pred_testing = FALSE
debug = F 

pheno_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_ischemic_phenotype.txt"
covar_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_ischemic_covariates.txt"
scores_file_vector <-fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_ischemic_file_paths.txt", header =F)$V1 
pgs_list1 <- fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_ischemic_PGS_list.txt", header = F)
trait_specific_score_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_ischemic_PGS_list.txt"


custom_pgs_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_ischemic_PGS_list.txt"
out_path <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/test_EUR_inc_ischemic_"
num_pgs <- nrow(pgs_list1)
pheno_name <- "inc_ischemic"

treatment <- NA
categorical_vars <- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd","prev_anycvd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")

categorical_vars<- setdiff(categorical_vars, treatment)
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

outcomes <- c("inc_anycmd", "inc_hyperten", "inc_diabetes", "inc_dyslipid", "inc_anycvd", "inc_seriouscvd", "inc_arrhythmia", "inc_hf", "inc_ischemic", "inc_stroke", "inc_vte", "inc_cvddeath")
cat_def <- c("M","M", "M","M", "V","V","V","V", "V", "V","V", "V")
dict <-setNames(cat_def, outcomes)



if(dict[[pheno_name]]=="M"){
  categorical_vars <- categorical_vars[categorical_vars !="prev_anycmd"]
}else{
  categorical_vars <-  categorical_vars[categorical_vars !="prev_anycvd"]
}

covariates <- c(categorical_vars, continuous_vars)

pheno_file <- pheno_file
covariate_file <- covar_file
score_files_list = scores_file_vector 
out <- out_path
IID_pheno = "IID" 
covar_list = covariates
cat_covar_list = categorical_vars
ncores = 1
is_extract_adjSNPeff = TRUE
power_thres_list = c(0.95) 
pval_thres_list = (.05/7)
nfold_cv = 3

```

```{r intro_chunk}
basic_data = fread(covariate_file)
basic_data = basic_data[,c(IID_pheno, covar_list)]
if (length(cat_covar_list)>0) {

	writeLines("Generating dummy variables from categorical variables")
	basic_data_list = make_dummy_columns(basic_data, cat_covar_list)

	covar_list = covar_list[-match(cat_covar_list, covar_list)]
	covar_list = c(covar_list, basic_data_list[[2]])

	basic_data = basic_data_list[[1]]
}

writeLines("--- Reading all polygenic risk scores ---")
writeLines("Please make sure the column names for the scores ends with _SUM")

all_scores = NULL

for (score_file_i in 1:length(score_files_list)) {

	print(score_file_i)
	score_file = score_files_list[score_file_i]
	dd = fread(score_file)
	idx = which(endsWith(colnames(dd), "_SUM") & colnames(dd)!="NAMED_ALLELE_DOSAGE_SUM")
	idx2 = which(colnames(dd)=="IID")

	dd_sub = dd[,c(idx2,idx)]
	print(dim(dd_sub))
	if (is.null(all_scores)) {
		all_scores = dd_sub
	} else
		all_scores = merge(all_scores, dd_sub, by="IID")
}

colnames(all_scores)[2:ncol(all_scores)] = substring(colnames(all_scores)[2:ncol(all_scores)], 1, nchar(colnames(all_scores)[2:ncol(all_scores)])-4)

if (!is.null(allPGS_list)) {
	pgs_extract = intersect(colnames(all_scores[2:ncol(all_scores)]), allPGS_list)
	all_scores = all_scores[,c(1,match(pgs_extract, colnames(all_scores)))]
}

score_names = colnames(all_scores)[2:ncol(all_scores)]

pgs_list = NULL
for (ff_i in 1:length(trait_specific_score_file)) {
	writeLines(paste0("Reading: ", trait_specific_score_file[ff_i]))
	pgs_list_tmp = fread(trait_specific_score_file[ff_i], header=F)[,1]
	pgs_list = c(pgs_list, pgs_list_tmp)
}
pgs_list = intersect(pgs_list, colnames(all_scores))

pheno = fread(pheno_file)
idx = which(colnames(pheno) %in% c(IID_pheno, pheno_name))
pheno = pheno[,idx]
colnames(pheno) = c("IID", "trait")

writeLines("--- Merging Phenotype and PRS files ---")

pheno_prs = merge(pheno, all_scores, by="IID")
pheno_prs_cov = merge(pheno_prs, basic_data, by.x="IID", by.y=IID_pheno)
pheno_prs_cov = pheno_prs_cov[which(!is.na(pheno_prs_cov$trait)),]

#######################

irnt = function(x) return(qnorm((rank(x,na.last="keep")-0.5)/sum(!is.na(x))))
rr = function(x,d=3) round(x,d)

#######################

null_train_size_list = F
if (is.null(train_size_list)) {
	writeLines("--- Selecting 80% as training data ---")
	train_size_list = floor(0.8*nrow(pheno_prs_cov))
	null_train_size_list = T
} else {
	writeLines("Using custom data split!")
}

out_save = out
```


```{r original_code}
for (train_size in train_size_list) {
  
    ################################ Set up Initial Data #################################
		train_size = train_size_list[1]
		writeLines(paste("--- Using ", train_size, " individuals for training sample ---"))
		if (!null_train_size_list) out = paste0(out_save, "_train.", train_size)
		if (isbinary) fwrite(as.data.frame(table(pheno_prs_cov$trait)), paste0(out, "_case_counts.txt"), row.names=F, sep="\t", quote=F)
		set.seed(1)
		train_idx = sample(1:nrow(pheno_prs_cov), train_size)
		train_df = pheno_prs_cov[train_idx,]
		test_df = pheno_prs_cov[-train_idx,]

		cc = apply(train_df[,covar_list], 2, function(x) length(unique(x)))
		covar_list1 = covar_list[-which(cc<=5)]
		for (i in covar_list1) train_df[i] = as.numeric(scale(train_df[i]))
		for (i in covar_list1) test_df[i] = as.numeric(scale(test_df[i]))

		fwrite(train_df[,c(1,2)], paste0(out, "_train_df.txt"), row.names=F, quote=F, sep="\t")
		fwrite(test_df[,c(1,2)], paste0(out, "_test_df.txt"), row.names=F, quote=F, sep="\t")

		################################ training #################################

		writeLines("--- Evaluating PRS in training set ---")
		
		if (!read_pred_training | is.null(training_result_file)) {
			training_file = paste0(out, "_train_allPRS.txt")
			read_pred_training_1 = (read_pred_training & file.exists(training_file))
		} else {
			training_file = training_result_file
			read_pred_training_1 = read_pred_training & all(file.exists(training_file))
			if (!read_pred_training_1) {
				writeLines("Declared:")
				writeLines(paste0("read_pred_training = ", read_pred_training))
				writeLines(paste0("training_result_file = ", paste(training_result_file, collapse=";")))
				stop("Reading existing training result but there is at least one missing training result file. Please check parameter all files available in <training_result_file> or set read_pred_testing=F and training_result_file=NULL")
			}
			writeLines(paste0("Reading all files in training_result_file"))
		}
		
		if (!read_pred_training_1) {

			sumscore = apply(as.data.frame(train_df[,score_names]), 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) train_df = train_df[,-match(names(idx), colnames(train_df))]

			pgs_list_all = colnames(train_df)
			pgs_list_all = pgs_list_all[which(pgs_list_all %in% colnames(all_scores)[2:ncol(all_scores)])]
			pred_acc_train_allPGS_summary = eval_multiple_PRS(train_df, pgs_list_all, covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)
			head(pred_acc_train_allPGS_summary)

			fwrite(pred_acc_train_allPGS_summary, training_file, row.names=F, sep="\t", quote=F)
		} else {
			pred_acc_train_allPGS_summary = NULL
			for (file_i in 1:length(training_file)) {
				writeLines(paste0("Reading training file: ", training_file[file_i]))
				pred_acc_train_allPGS_summary_tmp = fread(training_file[file_i])
				pred_acc_train_allPGS_summary = rbind(pred_acc_train_allPGS_summary, pred_acc_train_allPGS_summary_tmp)
			}
			pred_acc_train_allPGS_summary = pred_acc_train_allPGS_summary[!duplicated(pred_acc_train_allPGS_summary$pgs),]
		}

		pred_acc_train_allPGS_summary = as.data.frame(pred_acc_train_allPGS_summary)

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary
		pred_acc_train_trait_summary = pred_acc_train_trait_summary[order(as.numeric(pred_acc_train_trait_summary$pval), decreasing=F),]
		head(pred_acc_train_trait_summary)

		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary1[order(as.numeric(pred_acc_train_allPGS_summary1$R2), decreasing=T),]
		bestPRS = pred_acc_train_allPGS_summary1[1,1]
		writeLines(paste0("The best single trait-specific score in the training set is ", bestPRS))
		bestPRS_acc = eval_single_PRS(test_df, pheno = "trait", prs_name=bestPRS, covar_list=covar_list, liabilityR2, alpha=0.05, isbinary=isbinary)

		fwrite(bestPRS_acc, paste0(out, "_best_acc.txt"), row.names=F, sep="\t", quote=F)

		################################ testing #################################

		writeLines("--- Evaluating PRS in testing set ---")
		testing_file = paste0(out, "_test_allPRS.txt")
		read_pred_testing_1 = (read_pred_testing & file.exists(testing_file))
		if (!read_pred_testing_1) {
			sumscore = apply(test_df[,score_names], 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) test_df = test_df[,-match(names(idx), colnames(test_df))]

			pred_acc_test_trait = eval_multiple_PRS(test_df, pgs_list,  covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)

			pred_acc_test_trait_summary = pred_acc_test_trait
			pred_acc_test_trait_summary = pred_acc_test_trait_summary[order(as.numeric(pred_acc_test_trait_summary$pval), decreasing=F),]
			print(head(pred_acc_test_trait_summary))

			fwrite(pred_acc_test_trait_summary, testing_file, row.names=F, sep="\t", quote=F)
		} else {
			writeLines(paste0("Reading testing file: ", testing_file))
			pred_acc_test_trait_summary = fread(testing_file)
		}

		###########################################################################
		
		if (isbinary) {

			x_train = (train_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_train = as.vector(train_df$trait)
			train_data = data.frame(x_train,trait=y_train)

			x_test = (test_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_test = as.vector(test_df$trait)
			test_data = data.frame(x_test,trait=y_test)

			train_tmp = train_data[,c("trait", c(bestPRS, covar_list))]
			train_tmp$trait = as.factor(train_tmp$trait)
			
			formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, collapse="+")))
			formula_bestPRS = as.formula(paste0("trait ~ scale(", bestPRS, ")+", paste0(covar_list, collapse="+")))
			
			ctrl = trainControl(
				method = "repeatedcv",
				allowParallel = TRUE,
				number = nfold_cv,
				returnData = FALSE,
				trim = T,
				verboseIter = T)

			cl = makePSOCKcluster(ncores)
			cl = makeCluster(ncores)
			registerDoParallel(cl)

			set.seed(123)
			model_prsmix_null = train(
				formula_null, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)
			
			set.seed(123)
			model_prsmix_bestPRS = train(
				formula_bestPRS, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)

			stopCluster(cl)

			test_data1 = test_data
			
			# --- Null model AUC + ROC ---
      test_pred = predict(model_prsmix_null, test_data1, type = "prob")[,2]
      roc_null <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_null)
      auc_out = data.frame(method="null_model", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_NULL.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_null, paste0(out, "_ROC_null.rds"))
      
      # --- BestPRS model AUC + ROC ---
      test_pred = predict(model_prsmix_bestPRS, test_data1, type = "prob")[,2]
      roc_bestPRS <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_bestPRS)
      auc_out = data.frame(method="bestPGS", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_BestPGS.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_bestPRS, paste0(out, "_ROC_bestPRS.rds"))
			

		}

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		
		pred_acc_test_trait_summary_out = pred_acc_test_trait_summary

		writeLines("--- Iterating power and p-value parameters --- ")
		for (power_thres in power_thres_list)
			for (pval_thres in pval_thres_list) {
				
				writeLines(paste0("P = ", pval_thres))
				writeLines(paste0("Power = ", power_thres))
				writeLines("PRSmix:")

				topprs = pred_acc_train_trait_summary #%>%
					#filter(pval <= pval_thres & power >= power_thres)
				head(topprs)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))
				
				start_time = Sys.time()
				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix")
					ww_raw = 1
				} else {

					if (!isbinary) {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])					
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]

						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}
						
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)

						res_lm1 = eval_single_PRS(test_df1, pheno = "trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

					} else {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])						
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))
						
						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww_raw = ww

							# print(ww)
							# print(sd_train)
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww[which(!names(ww) %in% covar_list)]
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}

							if (debug) {
								print(ww)
							}
							# print(ww)
							test_data1 = test_data
							test_data1[,topprs] = as.numeric(scale(test_data[,topprs]))

							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmix", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmix.txt"), row.names=F, sep="\t", quote=F)
							
						}

						# print(head(test_data))	
						# print(head(test_df))	
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						
						res_lm1 = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

						
						############## OR ###################
						ff = paste0("trait ~ scale(newprs) + ", paste0(covar_list, collapse="+"))
						model1 = glm(ff, data=test_df1, family="binomial")
						model1s = summary(model1)
						mm = exp(model1s$coefficients[2,1])
						ll = exp(model1s$coefficients[2,1] - 1.97*model1s$coefficients[2,2])
						uu = exp(model1s$coefficients[2,1] + 1.97*model1s$coefficients[2,2])
						pval = format.pval(model1s$coefficients[2,4])
						print(paste0("OR = ", rr(mm), " (", rr(ll), "-", rr(uu), "); P-value=", pval))
						fwrite(data.frame(mm, ll, uu, pval), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_OR_PRSmix.txt"), row.names=F, sep="\t", quote=F)

						####################################

					}
					
					
					print(res_lm1)
					res_lm1_summary = res_lm1
					res_lm1_summary$pgs = "PRSmix"
					pred_acc_test_trait_summary_out = bind_rows(res_lm1, pred_acc_test_trait_summary)
					head(pred_acc_test_trait_summary_out)
					
					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmix.txt"), row.names=F, sep="\t", quote=F)
					
					prs_out = test_df1 %>%
						select(IID, newprs)
					colnames(prs_out) = c("IID", "prsmix")
					
					fwrite(prs_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmix.txt"), row.names=F, sep="\t", quote=F)
					
				}

				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]

				timedf = data.frame(pgs="PRSmix", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmix.txt"), row.names=F, sep="\t", quote=F)		

				############################

				writeLines("PRSmix+:")

				topprs = pred_acc_train_allPGS_summary# %>%
					#filter(pval <= pval_thres & power >= power_thres)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))

				start_time = Sys.time()

				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix+")
					ww_raw = 1
				} else {

					if (!isbinary) {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm

						nonzero_w = names(ww[which(ww!=0)])

					} else {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)
						
						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)
							
							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww==0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]							
							}
							
							test_data1 = test_data
							test_data1[,topprs] = scale(test_data[,topprs])
							
							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							roc_PRSMixPlus <- roc(test_data1$trait, test_pred)
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmixP", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", "newprs", covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm
						
						nonzero_w = names(ww[which(ww!=0)])
						
					}
					
					print(res_lm)
					pred_acc_test_trait_summary_out = bind_rows(res_lm, pred_acc_test_trait_summary_out)
					head(pred_acc_test_trait_summary_out)

					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

					prsmixplus = test_df1 %>% select(IID, newprs)

					fwrite(prsmixplus, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmixPlus.txt"), row.names=F, sep="\t", quote=F)

					fwrite(data.frame(topprs, ww), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)					
					
					if (is_extract_adjSNPeff) {
						mixing_weight_file = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt")
						outfile = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_adjSNPeff_PRSmixPlus.txt")
						extract_adjSNPeff(mixing_weight_file, original_beta_files_list, outfile)
					}
					
					fwrite(data.frame(topprs, ww_raw), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_raw_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
					
					##############################################
				}


				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]
				timedf = data.frame(pgs="PRSmix+", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
				

			}
	}
	writeLines("Finished")

```


```{r delong_test_isch}
# Compute CIs
auc_ci_null <- ci.auc(roc_null)
auc_ci_best <- ci.auc(roc_bestPRS)
auc_ci_prsmixplus <- ci.auc(roc_PRSMixPlus)


# Save AUC + CI
auc_out <- data.frame(
  method = c("null_model", "bestPGS", "PRSMix+"),
  auc = c(auc_ci_null[2], auc_ci_best[2], auc_ci_prsmixplus[2]),
  lowerCI = c(auc_ci_null[1], auc_ci_best[1], auc_ci_prsmixplus[1]),
  upperCI = c(auc_ci_null[3], auc_ci_best[3], auc_ci_prsmixplus[3])
)
#fwrite(auc_out, paste0(out, "_auc_summary.txt"), sep = "\t", quote = FALSE)

# 🧠 Compare AUCs using DeLong's test
auc_test1 <- roc.test(roc_null, roc_bestPRS, method = "delong")
auc_test2 <- roc.test(roc_null, roc_PRSMixPlus, method = "delong")

print(auc_test1)
print(auc_test2)
```

```{r isch_plot, fig.height = 6, fig.width = 6}
plot(
  roc_null,
  col = "#00A087FF", lwd = 2, legacy.axes = TRUE,
  main = "Ischemic Heart Disease" #,
  #print.auc = TRUE, print.auc.y = 0.4, print.auc.x = 0.6
)

plot(
  roc_bestPRS,
  col = "#4DBBD5FF", lwd = 2, add = TRUE #,
  #print.auc = TRUE, print.auc.y = 0.35, print.auc.x = 0.6
)

plot(
  roc_PRSMixPlus,
  col = "#E64B35FF", lwd = 2, add = TRUE#,
  #print.auc = TRUE, print.auc.y = 0.3, print.auc.x = 0.6
)

# Correct the legend: match length of colors/lwd with items
legend(
  "bottomright",
  legend = c(
    sprintf("Covariates Only (AUC = %.3f)", auc(roc_null)),
    sprintf("Covariates + Best PGS (AUC = %.3f)", auc(roc_bestPRS)),
    sprintf("Covariates + PRSMix+ (AUC = %.3f)", auc(roc_PRSMixPlus)),
    sprintf("BestPRS-Null: p = %.3g", auc_test1$p.value),
    sprintf("PRSMix+ - Null: p = %.3g", auc_test2$p.value)
  ),
  col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", NA, NA),
  lwd = c(2, 2, 2, NA, NA),
  bty = "n",
  text.col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", "black", "black")
)
# Save as a plot object
roc_plot_obj_ischemic <- recordPlot()


```

## Hypertension

```{r}


original_beta_files_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/janise_share/weights.txt"
isbinary <- TRUE
allPGS_list = NULL
metascore = NULL 
liabilityR2 = TRUE
train_size_list = NULL
training_result_file = NULL
read_pred_training = FALSE
read_pred_testing = FALSE
debug = F 

pheno_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_hyperten_phenotype.txt"
covar_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_hyperten_covariates.txt"
scores_file_vector <-fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_hyperten_file_paths.txt", header =F)$V1 
pgs_list1 <- fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_hyperten_PGS_list.txt", header = F)
trait_specific_score_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_hyperten_PGS_list.txt"


custom_pgs_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_hyperten_PGS_list.txt"
out_path <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/test_EUR_inc_hyperten_"
num_pgs <- nrow(pgs_list1)
pheno_name <- "inc_hyperten"

treatment <- NA
categorical_vars <- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd","prev_anycvd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")

categorical_vars<- setdiff(categorical_vars, treatment)
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

outcomes <- c("inc_anycmd", "inc_hyperten", "inc_diabetes", "inc_dyslipid", "inc_anycvd", "inc_seriouscvd", "inc_arrhythmia", "inc_hf", "inc_hyperten", "inc_stroke", "inc_vte", "inc_cvddeath")
cat_def <- c("M","M", "M","M", "V","V","V","V", "V", "V","V", "V")
dict <-setNames(cat_def, outcomes)



if(dict[[pheno_name]]=="M"){
  categorical_vars <- categorical_vars[categorical_vars !="prev_anycmd"]
}else{
  categorical_vars <-  categorical_vars[categorical_vars !="prev_anycvd"]
}

covariates <- c(categorical_vars, continuous_vars)

pheno_file <- pheno_file
covariate_file <- covar_file
score_files_list = scores_file_vector 
out <- out_path
IID_pheno = "IID" 
covar_list = covariates
cat_covar_list = categorical_vars
ncores = 1
is_extract_adjSNPeff = F
power_thres_list = c(0.95) 
pval_thres_list = (.05/7)
nfold_cv = 3
```

```{r intro_chunk_htn}
basic_data = fread(covariate_file)
basic_data = basic_data[,c(IID_pheno, covar_list)]
if (length(cat_covar_list)>0) {

	writeLines("Generating dummy variables from categorical variables")
	basic_data_list = make_dummy_columns(basic_data, cat_covar_list)

	covar_list = covar_list[-match(cat_covar_list, covar_list)]
	covar_list = c(covar_list, basic_data_list[[2]])

	basic_data = basic_data_list[[1]]
}

writeLines("--- Reading all polygenic risk scores ---")
writeLines("Please make sure the column names for the scores ends with _SUM")

all_scores = NULL

for (score_file_i in 1:length(score_files_list)) {

	print(score_file_i)
	score_file = score_files_list[score_file_i]
	dd = fread(score_file)
	idx = which(endsWith(colnames(dd), "_SUM") & colnames(dd)!="NAMED_ALLELE_DOSAGE_SUM")
	idx2 = which(colnames(dd)=="IID")

	dd_sub = dd[,c(idx2,idx)]
	print(dim(dd_sub))
	if (is.null(all_scores)) {
		all_scores = dd_sub
	} else
		all_scores = merge(all_scores, dd_sub, by="IID")
}

colnames(all_scores)[2:ncol(all_scores)] = substring(colnames(all_scores)[2:ncol(all_scores)], 1, nchar(colnames(all_scores)[2:ncol(all_scores)])-4)

if (!is.null(allPGS_list)) {
	pgs_extract = intersect(colnames(all_scores[2:ncol(all_scores)]), allPGS_list)
	all_scores = all_scores[,c(1,match(pgs_extract, colnames(all_scores)))]
}

score_names = colnames(all_scores)[2:ncol(all_scores)]

pgs_list = NULL
for (ff_i in 1:length(trait_specific_score_file)) {
	writeLines(paste0("Reading: ", trait_specific_score_file[ff_i]))
	pgs_list_tmp = fread(trait_specific_score_file[ff_i], header=F)[,1]
	pgs_list = c(pgs_list, pgs_list_tmp)
}
pgs_list = intersect(pgs_list, colnames(all_scores))

pheno = fread(pheno_file)
idx = which(colnames(pheno) %in% c(IID_pheno, pheno_name))
pheno = pheno[,idx]
colnames(pheno) = c("IID", "trait")

writeLines("--- Merging Phenotype and PRS files ---")

pheno_prs = merge(pheno, all_scores, by="IID")
pheno_prs_cov = merge(pheno_prs, basic_data, by.x="IID", by.y=IID_pheno)
pheno_prs_cov = pheno_prs_cov[which(!is.na(pheno_prs_cov$trait)),]

#######################

irnt = function(x) return(qnorm((rank(x,na.last="keep")-0.5)/sum(!is.na(x))))
rr = function(x,d=3) round(x,d)

#######################

null_train_size_list = F
if (is.null(train_size_list)) {
	writeLines("--- Selecting 80% as training data ---")
	train_size_list = floor(0.8*nrow(pheno_prs_cov))
	null_train_size_list = T
} else {
	writeLines("Using custom data split!")
}

out_save = out
```


```{r original_code_htn}
for (train_size in train_size_list) {
  
    ################################ Set up Initial Data #################################
		train_size = train_size_list[1]
		writeLines(paste("--- Using ", train_size, " individuals for training sample ---"))
		if (!null_train_size_list) out = paste0(out_save, "_train.", train_size)
		if (isbinary) fwrite(as.data.frame(table(pheno_prs_cov$trait)), paste0(out, "_case_counts.txt"), row.names=F, sep="\t", quote=F)
		set.seed(1)
		train_idx = sample(1:nrow(pheno_prs_cov), train_size)
		train_df = pheno_prs_cov[train_idx,]
		test_df = pheno_prs_cov[-train_idx,]

		cc = apply(train_df[,covar_list], 2, function(x) length(unique(x)))
		covar_list1 = covar_list[-which(cc<=5)]
		for (i in covar_list1) train_df[i] = as.numeric(scale(train_df[i]))
		for (i in covar_list1) test_df[i] = as.numeric(scale(test_df[i]))

		fwrite(train_df[,c(1,2)], paste0(out, "_train_df.txt"), row.names=F, quote=F, sep="\t")
		fwrite(test_df[,c(1,2)], paste0(out, "_test_df.txt"), row.names=F, quote=F, sep="\t")

		################################ training #################################

		writeLines("--- Evaluating PRS in training set ---")
		
		if (!read_pred_training | is.null(training_result_file)) {
			training_file = paste0(out, "_train_allPRS.txt")
			read_pred_training_1 = (read_pred_training & file.exists(training_file))
		} else {
			training_file = training_result_file
			read_pred_training_1 = read_pred_training & all(file.exists(training_file))
			if (!read_pred_training_1) {
				writeLines("Declared:")
				writeLines(paste0("read_pred_training = ", read_pred_training))
				writeLines(paste0("training_result_file = ", paste(training_result_file, collapse=";")))
				stop("Reading existing training result but there is at least one missing training result file. Please check parameter all files available in <training_result_file> or set read_pred_testing=F and training_result_file=NULL")
			}
			writeLines(paste0("Reading all files in training_result_file"))
		}
		
		if (!read_pred_training_1) {

			sumscore = apply(as.data.frame(train_df[,score_names]), 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) train_df = train_df[,-match(names(idx), colnames(train_df))]

			pgs_list_all = colnames(train_df)
			pgs_list_all = pgs_list_all[which(pgs_list_all %in% colnames(all_scores)[2:ncol(all_scores)])]
			pred_acc_train_allPGS_summary = eval_multiple_PRS(train_df, pgs_list_all, covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)
			head(pred_acc_train_allPGS_summary)

			fwrite(pred_acc_train_allPGS_summary, training_file, row.names=F, sep="\t", quote=F)
		} else {
			pred_acc_train_allPGS_summary = NULL
			for (file_i in 1:length(training_file)) {
				writeLines(paste0("Reading training file: ", training_file[file_i]))
				pred_acc_train_allPGS_summary_tmp = fread(training_file[file_i])
				pred_acc_train_allPGS_summary = rbind(pred_acc_train_allPGS_summary, pred_acc_train_allPGS_summary_tmp)
			}
			pred_acc_train_allPGS_summary = pred_acc_train_allPGS_summary[!duplicated(pred_acc_train_allPGS_summary$pgs),]
		}

		pred_acc_train_allPGS_summary = as.data.frame(pred_acc_train_allPGS_summary)

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary
		pred_acc_train_trait_summary = pred_acc_train_trait_summary[order(as.numeric(pred_acc_train_trait_summary$pval), decreasing=F),]
		head(pred_acc_train_trait_summary)

		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary1[order(as.numeric(pred_acc_train_allPGS_summary1$R2), decreasing=T),]
		bestPRS = pred_acc_train_allPGS_summary1[1,1]
		writeLines(paste0("The best single trait-specific score in the training set is ", bestPRS))
		bestPRS_acc = eval_single_PRS(test_df, pheno = "trait", prs_name=bestPRS, covar_list=covar_list, liabilityR2, alpha=0.05, isbinary=isbinary)

		fwrite(bestPRS_acc, paste0(out, "_best_acc.txt"), row.names=F, sep="\t", quote=F)

		################################ testing #################################

		writeLines("--- Evaluating PRS in testing set ---")
		testing_file = paste0(out, "_test_allPRS.txt")
		read_pred_testing_1 = (read_pred_testing & file.exists(testing_file))
		if (!read_pred_testing_1) {
			sumscore = apply(test_df[,score_names], 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) test_df = test_df[,-match(names(idx), colnames(test_df))]

			pred_acc_test_trait = eval_multiple_PRS(test_df, pgs_list,  covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)

			pred_acc_test_trait_summary = pred_acc_test_trait
			pred_acc_test_trait_summary = pred_acc_test_trait_summary[order(as.numeric(pred_acc_test_trait_summary$pval), decreasing=F),]
			print(head(pred_acc_test_trait_summary))

			fwrite(pred_acc_test_trait_summary, testing_file, row.names=F, sep="\t", quote=F)
		} else {
			writeLines(paste0("Reading testing file: ", testing_file))
			pred_acc_test_trait_summary = fread(testing_file)
		}

		###########################################################################
		
		if (isbinary) {

			x_train = (train_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_train = as.vector(train_df$trait)
			train_data = data.frame(x_train,trait=y_train)

			x_test = (test_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_test = as.vector(test_df$trait)
			test_data = data.frame(x_test,trait=y_test)

			train_tmp = train_data[,c("trait", c(bestPRS, covar_list))]
			train_tmp$trait = as.factor(train_tmp$trait)
			
			formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, collapse="+")))
			formula_bestPRS = as.formula(paste0("trait ~ scale(", bestPRS, ")+", paste0(covar_list, collapse="+")))
			
			ctrl = trainControl(
				method = "repeatedcv",
				allowParallel = TRUE,
				number = nfold_cv,
				returnData = FALSE,
				trim = T,
				verboseIter = T)

			cl = makePSOCKcluster(ncores)
			cl = makeCluster(ncores)
			registerDoParallel(cl)

			set.seed(123)
			model_prsmix_null = train(
				formula_null, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)
			
			set.seed(123)
			model_prsmix_bestPRS = train(
				formula_bestPRS, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)

			stopCluster(cl)

			test_data1 = test_data
			
			# --- Null model AUC + ROC ---
      test_pred = predict(model_prsmix_null, test_data1, type = "prob")[,2]
      roc_null <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_null)
      auc_out = data.frame(method="null_model", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_NULL.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_null, paste0(out, "_ROC_null.rds"))
      
      # --- BestPRS model AUC + ROC ---
      test_pred = predict(model_prsmix_bestPRS, test_data1, type = "prob")[,2]
      roc_bestPRS <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_bestPRS)
      auc_out = data.frame(method="bestPGS", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_BestPGS.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_bestPRS, paste0(out, "_ROC_bestPRS.rds"))
			

		}

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		
		pred_acc_test_trait_summary_out = pred_acc_test_trait_summary

		writeLines("--- Iterating power and p-value parameters --- ")
		for (power_thres in power_thres_list)
			for (pval_thres in pval_thres_list) {
				
				writeLines(paste0("P = ", pval_thres))
				writeLines(paste0("Power = ", power_thres))
				writeLines("PRSmix:")

				topprs = pred_acc_train_trait_summary #%>%
					#filter(pval <= pval_thres & power >= power_thres)
				head(topprs)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))
				
				start_time = Sys.time()
				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix")
					ww_raw = 1
				} else {

					if (!isbinary) {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])					
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]

						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}
						
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)

						res_lm1 = eval_single_PRS(test_df1, pheno = "trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

					} else {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])						
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))
						
						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww_raw = ww

							# print(ww)
							# print(sd_train)
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww[which(!names(ww) %in% covar_list)]
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}

							if (debug) {
								print(ww)
							}
							# print(ww)
							test_data1 = test_data
							test_data1[,topprs] = as.numeric(scale(test_data[,topprs]))

							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmix", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmix.txt"), row.names=F, sep="\t", quote=F)
							
						}

						# print(head(test_data))	
						# print(head(test_df))	
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						
						res_lm1 = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

						
						############## OR ###################
						ff = paste0("trait ~ scale(newprs) + ", paste0(covar_list, collapse="+"))
						model1 = glm(ff, data=test_df1, family="binomial")
						model1s = summary(model1)
						mm = exp(model1s$coefficients[2,1])
						ll = exp(model1s$coefficients[2,1] - 1.97*model1s$coefficients[2,2])
						uu = exp(model1s$coefficients[2,1] + 1.97*model1s$coefficients[2,2])
						pval = format.pval(model1s$coefficients[2,4])
						print(paste0("OR = ", rr(mm), " (", rr(ll), "-", rr(uu), "); P-value=", pval))
						fwrite(data.frame(mm, ll, uu, pval), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_OR_PRSmix.txt"), row.names=F, sep="\t", quote=F)

						####################################

					}
					
					
					print(res_lm1)
					res_lm1_summary = res_lm1
					res_lm1_summary$pgs = "PRSmix"
					pred_acc_test_trait_summary_out = bind_rows(res_lm1, pred_acc_test_trait_summary)
					head(pred_acc_test_trait_summary_out)
					
					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmix.txt"), row.names=F, sep="\t", quote=F)
					
					prs_out = test_df1 %>%
						select(IID, newprs)
					colnames(prs_out) = c("IID", "prsmix")
					
					fwrite(prs_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmix.txt"), row.names=F, sep="\t", quote=F)
					
				}

				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]

				timedf = data.frame(pgs="PRSmix", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmix.txt"), row.names=F, sep="\t", quote=F)		

				############################

				writeLines("PRSmix+:")

				topprs = pred_acc_train_allPGS_summary #%>%
					#filter(pval <= pval_thres & power >= power_thres)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))

				start_time = Sys.time()

				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix+")
					ww_raw = 1
				} else {

					if (!isbinary) {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm

						nonzero_w = names(ww[which(ww!=0)])

					} else {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)
						
						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)
							
							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww==0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]							
							}
							
							test_data1 = test_data
							test_data1[,topprs] = scale(test_data[,topprs])
							
							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							roc_PRSMixPlus <- roc(test_data1$trait, test_pred)
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmixP", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", "newprs", covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm
						
						nonzero_w = names(ww[which(ww!=0)])
						
					}
					
					print(res_lm)
					pred_acc_test_trait_summary_out = bind_rows(res_lm, pred_acc_test_trait_summary_out)
					head(pred_acc_test_trait_summary_out)

					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

					prsmixplus = test_df1 %>% select(IID, newprs)

					fwrite(prsmixplus, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmixPlus.txt"), row.names=F, sep="\t", quote=F)

					fwrite(data.frame(topprs, ww), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)					
					
					if (is_extract_adjSNPeff) {
						mixing_weight_file = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt")
						outfile = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_adjSNPeff_PRSmixPlus.txt")
						extract_adjSNPeff(mixing_weight_file, original_beta_files_list, outfile)
					}
					
					fwrite(data.frame(topprs, ww_raw), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_raw_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
					
					##############################################
				}


				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]
				timedf = data.frame(pgs="PRSmix+", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
				

			}
	}
	writeLines("Finished")

```


```{r delong_test_hypertension}
# Compute CIs
auc_ci_null <- ci.auc(roc_null)
auc_ci_best <- ci.auc(roc_bestPRS)
auc_ci_prsmixplus <- ci.auc(roc_PRSMixPlus)


# Save AUC + CI
auc_out <- data.frame(
  method = c("null_model", "bestPGS", "PRSMix+"),
  auc = c(auc_ci_null[2], auc_ci_best[2], auc_ci_prsmixplus[2]),
  lowerCI = c(auc_ci_null[1], auc_ci_best[1], auc_ci_prsmixplus[1]),
  upperCI = c(auc_ci_null[3], auc_ci_best[3], auc_ci_prsmixplus[3])
)
#fwrite(auc_out, paste0(out, "_auc_summary.txt"), sep = "\t", quote = FALSE)

# 🧠 Compare AUCs using DeLong's test
auc_test1 <- roc.test(roc_null, roc_bestPRS, method = "delong")
auc_test2 <- roc.test(roc_null, roc_PRSMixPlus, method = "delong")

print(auc_test1)
print(auc_test2)
```

```{r hypertension_plot, fig.height = 6, fig.width = 6}
plot(
  roc_null,
  col = "#00A087FF", lwd = 2, legacy.axes = TRUE,
  main = "Hypertension" #,
  #print.auc = TRUE, print.auc.y = 0.4, print.auc.x = 0.6
)

plot(
  roc_bestPRS,
  col = "#4DBBD5FF", lwd = 2, add = TRUE #,
  #print.auc = TRUE, print.auc.y = 0.35, print.auc.x = 0.6
)

plot(
  roc_PRSMixPlus,
  col = "#E64B35FF", lwd = 2, add = TRUE#,
  #print.auc = TRUE, print.auc.y = 0.3, print.auc.x = 0.6
)

# Correct the legend: match length of colors/lwd with items
legend(
  "bottomright",
  legend = c(
    sprintf("Covariates Only (AUC = %.3f)", auc(roc_null)),
    sprintf("Covariates + Best PGS (AUC = %.3f)", auc(roc_bestPRS)),
    sprintf("Covariates + PRSMix+ (AUC = %.3f)", auc(roc_PRSMixPlus)),
    sprintf("BestPRS-Null: p = %.3g", auc_test1$p.value),
    sprintf("PRSMix+ - Null: p = %.3g", auc_test2$p.value)
  ),
  col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", NA, NA),
  lwd = c(2, 2, 2, NA, NA),
  bty = "n",
  text.col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", "black", "black")
)
# Save as a plot object
roc_plot_obj_htn <- recordPlot()


```

## Dyslipidemia


```{r}
original_beta_files_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/janise_share/weights.txt"
isbinary <- TRUE
allPGS_list = NULL
metascore = NULL 
liabilityR2 = TRUE
train_size_list = NULL
training_result_file = NULL
read_pred_training = FALSE
read_pred_testing = FALSE
debug = F 

pheno_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_dyslipid_phenotype.txt"
covar_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_dyslipid_covariates.txt"
scores_file_vector <-fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_dyslipid_file_paths.txt", header =F)$V1 
pgs_list1 <- fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_dyslipid_PGS_list.txt", header = F)
trait_specific_score_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_dyslipid_PGS_list.txt"


custom_pgs_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_dyslipid_PGS_list.txt"
out_path <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/test_EUR_inc_dyslipid_"
num_pgs <- nrow(pgs_list1)
pheno_name <- "inc_dyslipid"

treatment <- NA
categorical_vars <- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd","prev_anycvd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")

categorical_vars<- setdiff(categorical_vars, treatment)
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

outcomes <- c("inc_anycmd", "inc_hyperten", "inc_diabetes", "inc_dyslipid", "inc_anycvd", "inc_seriouscvd", "inc_arrhythmia", "inc_hf", "inc_ischemic", "inc_stroke", "inc_vte", "inc_cvddeath")
cat_def <- c("M","M", "M","M", "V","V","V","V", "V", "V","V", "V")
dict <-setNames(cat_def, outcomes)



if(dict[[pheno_name]]=="M"){
  categorical_vars <- categorical_vars[categorical_vars !="prev_anycmd"]
}else{
  categorical_vars <-  categorical_vars[categorical_vars !="prev_anycvd"]
}

covariates <- c(categorical_vars, continuous_vars)

allPGS_list = NULL
pheno_file <- pheno_file
covariate_file <- covar_file
score_files_list = scores_file_vector 
out <- out_path
IID_pheno = "IID" 
covar_list = covariates
cat_covar_list = categorical_vars
ncores = 1
is_extract_adjSNPeff = F
power_thres_list = c(0.95) 
pval_thres_list = (.05/7)
nfold_cv = 3
```

```{r intro_chunk_dyslipid}
basic_data = fread(covariate_file)
basic_data = basic_data[,c(IID_pheno, covar_list)]
if (length(cat_covar_list)>0) {

	writeLines("Generating dummy variables from categorical variables")
	basic_data_list = make_dummy_columns(basic_data, cat_covar_list)

	covar_list = covar_list[-match(cat_covar_list, covar_list)]
	covar_list = c(covar_list, basic_data_list[[2]])

	basic_data = basic_data_list[[1]]
}

writeLines("--- Reading all polygenic risk scores ---")
writeLines("Please make sure the column names for the scores ends with _SUM")

all_scores = NULL

for (score_file_i in 1:length(score_files_list)) {

	print(score_file_i)
	score_file = score_files_list[score_file_i]
	dd = fread(score_file)
	idx = which(endsWith(colnames(dd), "_SUM") & colnames(dd)!="NAMED_ALLELE_DOSAGE_SUM")
	idx2 = which(colnames(dd)=="IID")

	dd_sub = dd[,c(idx2,idx)]
	print(dim(dd_sub))
	if (is.null(all_scores)) {
		all_scores = dd_sub
	} else
		all_scores = merge(all_scores, dd_sub, by="IID")
}

colnames(all_scores)[2:ncol(all_scores)] = substring(colnames(all_scores)[2:ncol(all_scores)], 1, nchar(colnames(all_scores)[2:ncol(all_scores)])-4)

if (!is.null(allPGS_list)) {
	pgs_extract = intersect(colnames(all_scores[2:ncol(all_scores)]), allPGS_list)
	all_scores = all_scores[,c(1,match(pgs_extract, colnames(all_scores)))]
}

score_names = colnames(all_scores)[2:ncol(all_scores)]

pgs_list = NULL
for (ff_i in 1:length(trait_specific_score_file)) {
	writeLines(paste0("Reading: ", trait_specific_score_file[ff_i]))
	pgs_list_tmp = fread(trait_specific_score_file[ff_i], header=F)[,1]
	pgs_list = c(pgs_list, pgs_list_tmp)
}
pgs_list = intersect(pgs_list, colnames(all_scores))

pheno = fread(pheno_file)
idx = which(colnames(pheno) %in% c(IID_pheno, pheno_name))
pheno = pheno[,idx]
colnames(pheno) = c("IID", "trait")

writeLines("--- Merging Phenotype and PRS files ---")

pheno_prs = merge(pheno, all_scores, by="IID")
pheno_prs_cov = merge(pheno_prs, basic_data, by.x="IID", by.y=IID_pheno)
pheno_prs_cov = pheno_prs_cov[which(!is.na(pheno_prs_cov$trait)),]

#######################

irnt = function(x) return(qnorm((rank(x,na.last="keep")-0.5)/sum(!is.na(x))))
rr = function(x,d=3) round(x,d)

#######################

null_train_size_list = F
if (is.null(train_size_list)) {
	writeLines("--- Selecting 80% as training data ---")
	train_size_list = floor(0.8*nrow(pheno_prs_cov))
	null_train_size_list = T
} else {
	writeLines("Using custom data split!")
}

out_save = out
```


```{r original_code_dyslipid}
for (train_size in train_size_list) {
  
    ################################ Set up Initial Data #################################
		train_size = train_size_list[1]
		writeLines(paste("--- Using ", train_size, " individuals for training sample ---"))
		if (!null_train_size_list) out = paste0(out_save, "_train.", train_size)
		if (isbinary) fwrite(as.data.frame(table(pheno_prs_cov$trait)), paste0(out, "_case_counts.txt"), row.names=F, sep="\t", quote=F)
		set.seed(1)
		train_idx = sample(1:nrow(pheno_prs_cov), train_size)
		train_df = pheno_prs_cov[train_idx,]
		test_df = pheno_prs_cov[-train_idx,]

		cc = apply(train_df[,covar_list], 2, function(x) length(unique(x)))
		covar_list1 = covar_list[-which(cc<=5)]
		for (i in covar_list1) train_df[i] = as.numeric(scale(train_df[i]))
		for (i in covar_list1) test_df[i] = as.numeric(scale(test_df[i]))

		fwrite(train_df[,c(1,2)], paste0(out, "_train_df.txt"), row.names=F, quote=F, sep="\t")
		fwrite(test_df[,c(1,2)], paste0(out, "_test_df.txt"), row.names=F, quote=F, sep="\t")

		################################ training #################################

		writeLines("--- Evaluating PRS in training set ---")
		
		if (!read_pred_training | is.null(training_result_file)) {
			training_file = paste0(out, "_train_allPRS.txt")
			read_pred_training_1 = (read_pred_training & file.exists(training_file))
		} else {
			training_file = training_result_file
			read_pred_training_1 = read_pred_training & all(file.exists(training_file))
			if (!read_pred_training_1) {
				writeLines("Declared:")
				writeLines(paste0("read_pred_training = ", read_pred_training))
				writeLines(paste0("training_result_file = ", paste(training_result_file, collapse=";")))
				stop("Reading existing training result but there is at least one missing training result file. Please check parameter all files available in <training_result_file> or set read_pred_testing=F and training_result_file=NULL")
			}
			writeLines(paste0("Reading all files in training_result_file"))
		}
		
		if (!read_pred_training_1) {

			sumscore = apply(as.data.frame(train_df[,score_names]), 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) train_df = train_df[,-match(names(idx), colnames(train_df))]

			pgs_list_all = colnames(train_df)
			pgs_list_all = pgs_list_all[which(pgs_list_all %in% colnames(all_scores)[2:ncol(all_scores)])]
			pred_acc_train_allPGS_summary = eval_multiple_PRS(train_df, pgs_list_all, covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)
			head(pred_acc_train_allPGS_summary)

			fwrite(pred_acc_train_allPGS_summary, training_file, row.names=F, sep="\t", quote=F)
		} else {
			pred_acc_train_allPGS_summary = NULL
			for (file_i in 1:length(training_file)) {
				writeLines(paste0("Reading training file: ", training_file[file_i]))
				pred_acc_train_allPGS_summary_tmp = fread(training_file[file_i])
				pred_acc_train_allPGS_summary = rbind(pred_acc_train_allPGS_summary, pred_acc_train_allPGS_summary_tmp)
			}
			pred_acc_train_allPGS_summary = pred_acc_train_allPGS_summary[!duplicated(pred_acc_train_allPGS_summary$pgs),]
		}

		pred_acc_train_allPGS_summary = as.data.frame(pred_acc_train_allPGS_summary)

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary
		pred_acc_train_trait_summary = pred_acc_train_trait_summary[order(as.numeric(pred_acc_train_trait_summary$pval), decreasing=F),]
		head(pred_acc_train_trait_summary)

		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary1[order(as.numeric(pred_acc_train_allPGS_summary1$R2), decreasing=T),]
		bestPRS = pred_acc_train_allPGS_summary1[1,1]
		writeLines(paste0("The best single trait-specific score in the training set is ", bestPRS))
		bestPRS_acc = eval_single_PRS(test_df, pheno = "trait", prs_name=bestPRS, covar_list=covar_list, liabilityR2, alpha=0.05, isbinary=isbinary)

		fwrite(bestPRS_acc, paste0(out, "_best_acc.txt"), row.names=F, sep="\t", quote=F)

		################################ testing #################################

		writeLines("--- Evaluating PRS in testing set ---")
		testing_file = paste0(out, "_test_allPRS.txt")
		read_pred_testing_1 = (read_pred_testing & file.exists(testing_file))
		if (!read_pred_testing_1) {
			sumscore = apply(test_df[,score_names], 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) test_df = test_df[,-match(names(idx), colnames(test_df))]

			pred_acc_test_trait = eval_multiple_PRS(test_df, pgs_list,  covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)

			pred_acc_test_trait_summary = pred_acc_test_trait
			pred_acc_test_trait_summary = pred_acc_test_trait_summary[order(as.numeric(pred_acc_test_trait_summary$pval), decreasing=F),]
			print(head(pred_acc_test_trait_summary))

			fwrite(pred_acc_test_trait_summary, testing_file, row.names=F, sep="\t", quote=F)
		} else {
			writeLines(paste0("Reading testing file: ", testing_file))
			pred_acc_test_trait_summary = fread(testing_file)
		}

		###########################################################################
		
		if (isbinary) {

			x_train = (train_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_train = as.vector(train_df$trait)
			train_data = data.frame(x_train,trait=y_train)

			x_test = (test_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_test = as.vector(test_df$trait)
			test_data = data.frame(x_test,trait=y_test)

			train_tmp = train_data[,c("trait", c(bestPRS, covar_list))]
			train_tmp$trait = as.factor(train_tmp$trait)
			
			formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, collapse="+")))
			formula_bestPRS = as.formula(paste0("trait ~ scale(", bestPRS, ")+", paste0(covar_list, collapse="+")))
			
			ctrl = trainControl(
				method = "repeatedcv",
				allowParallel = TRUE,
				number = nfold_cv,
				returnData = FALSE,
				trim = T,
				verboseIter = T)

			cl = makePSOCKcluster(ncores)
			cl = makeCluster(ncores)
			registerDoParallel(cl)

			set.seed(123)
			model_prsmix_null = train(
				formula_null, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)
			
			set.seed(123)
			model_prsmix_bestPRS = train(
				formula_bestPRS, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)

			stopCluster(cl)

			test_data1 = test_data
			
			# --- Null model AUC + ROC ---
      test_pred = predict(model_prsmix_null, test_data1, type = "prob")[,2]
      roc_null <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_null)
      auc_out = data.frame(method="null_model", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_NULL.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_null, paste0(out, "_ROC_null.rds"))
      
      # --- BestPRS model AUC + ROC ---
      test_pred = predict(model_prsmix_bestPRS, test_data1, type = "prob")[,2]
      roc_bestPRS <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_bestPRS)
      auc_out = data.frame(method="bestPGS", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_BestPGS.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_bestPRS, paste0(out, "_ROC_bestPRS.rds"))
			

		}

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		
		pred_acc_test_trait_summary_out = pred_acc_test_trait_summary

		writeLines("--- Iterating power and p-value parameters --- ")
		for (power_thres in power_thres_list)
			for (pval_thres in pval_thres_list) {
				
				writeLines(paste0("P = ", pval_thres))
				writeLines(paste0("Power = ", power_thres))
				writeLines("PRSmix:")

				topprs = pred_acc_train_trait_summary %>%
					filter(pval <= pval_thres & power >= power_thres)
				head(topprs)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))
				
				start_time = Sys.time()
				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix")
					ww_raw = 1
				} else {

					if (!isbinary) {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])					
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]

						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}
						
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)

						res_lm1 = eval_single_PRS(test_df1, pheno = "trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

					} else {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])						
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))
						
						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww_raw = ww

							# print(ww)
							# print(sd_train)
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww[which(!names(ww) %in% covar_list)]
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}

							if (debug) {
								print(ww)
							}
							# print(ww)
							test_data1 = test_data
							test_data1[,topprs] = as.numeric(scale(test_data[,topprs]))

							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmix", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmix.txt"), row.names=F, sep="\t", quote=F)
							
						}

						# print(head(test_data))	
						# print(head(test_df))	
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						
						res_lm1 = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

						
						############## OR ###################
						ff = paste0("trait ~ scale(newprs) + ", paste0(covar_list, collapse="+"))
						model1 = glm(ff, data=test_df1, family="binomial")
						model1s = summary(model1)
						mm = exp(model1s$coefficients[2,1])
						ll = exp(model1s$coefficients[2,1] - 1.97*model1s$coefficients[2,2])
						uu = exp(model1s$coefficients[2,1] + 1.97*model1s$coefficients[2,2])
						pval = format.pval(model1s$coefficients[2,4])
						print(paste0("OR = ", rr(mm), " (", rr(ll), "-", rr(uu), "); P-value=", pval))
						fwrite(data.frame(mm, ll, uu, pval), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_OR_PRSmix.txt"), row.names=F, sep="\t", quote=F)

						####################################

					}
					
					
					print(res_lm1)
					res_lm1_summary = res_lm1
					res_lm1_summary$pgs = "PRSmix"
					pred_acc_test_trait_summary_out = bind_rows(res_lm1, pred_acc_test_trait_summary)
					head(pred_acc_test_trait_summary_out)
					
					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmix.txt"), row.names=F, sep="\t", quote=F)
					
					prs_out = test_df1 %>%
						select(IID, newprs)
					colnames(prs_out) = c("IID", "prsmix")
					
					fwrite(prs_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmix.txt"), row.names=F, sep="\t", quote=F)
					
				}

				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]

				timedf = data.frame(pgs="PRSmix", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmix.txt"), row.names=F, sep="\t", quote=F)		

				############################

				writeLines("PRSmix+:")

				topprs = pred_acc_train_allPGS_summary  %>%
					filter(pval <= pval_thres & power >= power_thres)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))

				start_time = Sys.time()

				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix+")
					ww_raw = 1
				} else {

					if (!isbinary) {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm

						nonzero_w = names(ww[which(ww!=0)])

					} else {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)
						
						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)
							
							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww==0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]							
							}
							
							test_data1 = test_data
							test_data1[,topprs] = scale(test_data[,topprs])
							
							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							roc_PRSMixPlus <- roc(test_data1$trait, test_pred)
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmixP", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", "newprs", covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm
						
						nonzero_w = names(ww[which(ww!=0)])
						
					}
					
					print(res_lm)
					pred_acc_test_trait_summary_out = bind_rows(res_lm, pred_acc_test_trait_summary_out)
					head(pred_acc_test_trait_summary_out)

					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

					prsmixplus = test_df1 %>% select(IID, newprs)

					fwrite(prsmixplus, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmixPlus.txt"), row.names=F, sep="\t", quote=F)

					fwrite(data.frame(topprs, ww), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)					
					
					if (is_extract_adjSNPeff) {
						mixing_weight_file = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt")
						outfile = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_adjSNPeff_PRSmixPlus.txt")
						extract_adjSNPeff(mixing_weight_file, original_beta_files_list, outfile)
					}
					
					fwrite(data.frame(topprs, ww_raw), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_raw_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
					
					##############################################
				}


				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]
				timedf = data.frame(pgs="PRSmix+", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
				

			}
	}
	writeLines("Finished")

```


```{r delong_test_dyslipid}
# Compute CIs
auc_ci_null <- ci.auc(roc_null)
auc_ci_best <- ci.auc(roc_bestPRS)
auc_ci_prsmixplus <- ci.auc(roc_PRSMixPlus)


# Save AUC + CI
auc_out <- data.frame(
  method = c("null_model", "bestPGS", "PRSMix+"),
  auc = c(auc_ci_null[2], auc_ci_best[2], auc_ci_prsmixplus[2]),
  lowerCI = c(auc_ci_null[1], auc_ci_best[1], auc_ci_prsmixplus[1]),
  upperCI = c(auc_ci_null[3], auc_ci_best[3], auc_ci_prsmixplus[3])
)
#fwrite(auc_out, paste0(out, "_auc_summary.txt"), sep = "\t", quote = FALSE)

# 🧠 Compare AUCs using DeLong's test
auc_test1 <- roc.test(roc_null, roc_bestPRS, method = "delong")
auc_test2 <- roc.test(roc_null, roc_PRSMixPlus, method = "delong")

print(auc_test1)
print(auc_test2)
```

```{r dyslipidemia_plot, fig.height = 6, fig.width = 6}
plot(
  roc_null,
  col = "#00A087FF", lwd = 2, legacy.axes = TRUE,
  main = "Dyslipidemia" #,
  #print.auc = TRUE, print.auc.y = 0.4, print.auc.x = 0.6
)

plot(
  roc_bestPRS,
  col = "#4DBBD5FF", lwd = 2, add = TRUE #,
  #print.auc = TRUE, print.auc.y = 0.35, print.auc.x = 0.6
)

plot(
  roc_PRSMixPlus,
  col = "#E64B35FF", lwd = 2, add = TRUE#,
  #print.auc = TRUE, print.auc.y = 0.3, print.auc.x = 0.6
)

# Correct the legend: match length of colors/lwd with items
legend(
  "bottomright",
  legend = c(
    sprintf("Covariates Only (AUC = %.3f)", auc(roc_null)),
    sprintf("Covariates + Best PGS (AUC = %.3f)", auc(roc_bestPRS)),
    sprintf("Covariates + PRSMix+ (AUC = %.3f)", auc(roc_PRSMixPlus)),
    sprintf("BestPRS-Null: p = %.3g", auc_test1$p.value),
    sprintf("PRSMix+ - Null: p = %.3g", auc_test2$p.value)
  ),
  col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", NA, NA),
  lwd = c(2, 2, 2, NA, NA),
  bty = "n",
  text.col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", "black", "black")
)
# Save as a plot object
roc_plot_obj_dyslipid <- recordPlot()


```

## Diabetes

```{r}
original_beta_files_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/janise_share/weights.txt"
isbinary <- TRUE
allPGS_list = NULL
metascore = NULL 
liabilityR2 = TRUE
train_size_list = NULL
training_result_file = NULL
read_pred_training = FALSE
read_pred_testing = FALSE
debug = F 

pheno_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_diabetes_phenotype.txt"
covar_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_diabetes_covariates.txt"
scores_file_vector <-fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_diabetes_file_paths.txt", header =F)$V1 
pgs_list1 <- fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_diabetes_PGS_list.txt", header = F)
trait_specific_score_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_diabetes_PGS_list.txt"


custom_pgs_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_diabetes_PGS_list.txt"
out_path <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/test_EUR_inc_diabetes_"
num_pgs <- nrow(pgs_list1)
pheno_name <- "inc_diabetes"

treatment <- NA
categorical_vars <- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd","prev_anycvd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")

categorical_vars<- setdiff(categorical_vars, treatment)
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

outcomes <- c("inc_anycmd", "inc_hyperten", "inc_diabetes", "inc_dyslipid", "inc_anycvd", "inc_seriouscvd", "inc_arrhythmia", "inc_hf", "inc_ischemic", "inc_stroke", "inc_vte", "inc_cvddeath")
cat_def <- c("M","M", "M","M", "V","V","V","V", "V", "V","V", "V")
dict <-setNames(cat_def, outcomes)



if(dict[[pheno_name]]=="M"){
  categorical_vars <- categorical_vars[categorical_vars !="prev_anycmd"]
}else{
  categorical_vars <-  categorical_vars[categorical_vars !="prev_anycvd"]
}

covariates <- c(categorical_vars, continuous_vars)

pheno_file <- pheno_file
covariate_file <- covar_file
score_files_list = scores_file_vector 
out <- out_path
IID_pheno = "IID" 
covar_list = covariates
cat_covar_list = categorical_vars
ncores = 1
is_extract_adjSNPeff = F
power_thres_list = c(0.95) 
pval_thres_list = .05
nfold_cv = 3
```

```{r intro_chunk_diabetes}
basic_data = fread(covariate_file)
basic_data = basic_data[,c(IID_pheno, covar_list)]
if (length(cat_covar_list)>0) {

	writeLines("Generating dummy variables from categorical variables")
	basic_data_list = make_dummy_columns(basic_data, cat_covar_list)

	covar_list = covar_list[-match(cat_covar_list, covar_list)]
	covar_list = c(covar_list, basic_data_list[[2]])

	basic_data = basic_data_list[[1]]
}

writeLines("--- Reading all polygenic risk scores ---")
writeLines("Please make sure the column names for the scores ends with _SUM")

all_scores = NULL

for (score_file_i in 1:length(score_files_list)) {

	print(score_file_i)
	score_file = score_files_list[score_file_i]
	dd = fread(score_file)
	idx = which(endsWith(colnames(dd), "_SUM") & colnames(dd)!="NAMED_ALLELE_DOSAGE_SUM")
	idx2 = which(colnames(dd)=="IID")

	dd_sub = dd[,c(idx2,idx)]
	print(dim(dd_sub))
	if (is.null(all_scores)) {
		all_scores = dd_sub
	} else
		all_scores = merge(all_scores, dd_sub, by="IID")
}

colnames(all_scores)[2:ncol(all_scores)] = substring(colnames(all_scores)[2:ncol(all_scores)], 1, nchar(colnames(all_scores)[2:ncol(all_scores)])-4)

if (!is.null(allPGS_list)) {
	pgs_extract = intersect(colnames(all_scores[2:ncol(all_scores)]), allPGS_list)
	all_scores = all_scores[,c(1,match(pgs_extract, colnames(all_scores)))]
}

score_names = colnames(all_scores)[2:ncol(all_scores)]

pgs_list = NULL
for (ff_i in 1:length(trait_specific_score_file)) {
	writeLines(paste0("Reading: ", trait_specific_score_file[ff_i]))
	pgs_list_tmp = fread(trait_specific_score_file[ff_i], header=F)[,1]
	pgs_list = c(pgs_list, pgs_list_tmp)
}
pgs_list = intersect(pgs_list, colnames(all_scores))

pheno = fread(pheno_file)
idx = which(colnames(pheno) %in% c(IID_pheno, pheno_name))
pheno = pheno[,idx]
colnames(pheno) = c("IID", "trait")

writeLines("--- Merging Phenotype and PRS files ---")

pheno_prs = merge(pheno, all_scores, by="IID")
pheno_prs_cov = merge(pheno_prs, basic_data, by.x="IID", by.y=IID_pheno)
pheno_prs_cov = pheno_prs_cov[which(!is.na(pheno_prs_cov$trait)),]

#######################

irnt = function(x) return(qnorm((rank(x,na.last="keep")-0.5)/sum(!is.na(x))))
rr = function(x,d=3) round(x,d)

#######################

null_train_size_list = F
if (is.null(train_size_list)) {
	writeLines("--- Selecting 80% as training data ---")
	train_size_list = floor(0.8*nrow(pheno_prs_cov))
	null_train_size_list = T
} else {
	writeLines("Using custom data split!")
}

out_save = out
```


```{r original_code_diabetes}
for (train_size in train_size_list) {
  
    ################################ Set up Initial Data #################################
		train_size = train_size_list[1]
		writeLines(paste("--- Using ", train_size, " individuals for training sample ---"))
		if (!null_train_size_list) out = paste0(out_save, "_train.", train_size)
		if (isbinary) fwrite(as.data.frame(table(pheno_prs_cov$trait)), paste0(out, "_case_counts.txt"), row.names=F, sep="\t", quote=F)
		set.seed(1)
		train_idx = sample(1:nrow(pheno_prs_cov), train_size)
		train_df = pheno_prs_cov[train_idx,]
		test_df = pheno_prs_cov[-train_idx,]

		cc = apply(train_df[,covar_list], 2, function(x) length(unique(x)))
		covar_list1 = covar_list[-which(cc<=5)]
		for (i in covar_list1) train_df[i] = as.numeric(scale(train_df[i]))
		for (i in covar_list1) test_df[i] = as.numeric(scale(test_df[i]))

		fwrite(train_df[,c(1,2)], paste0(out, "_train_df.txt"), row.names=F, quote=F, sep="\t")
		fwrite(test_df[,c(1,2)], paste0(out, "_test_df.txt"), row.names=F, quote=F, sep="\t")

		################################ training #################################

		writeLines("--- Evaluating PRS in training set ---")
		
		if (!read_pred_training | is.null(training_result_file)) {
			training_file = paste0(out, "_train_allPRS.txt")
			read_pred_training_1 = (read_pred_training & file.exists(training_file))
		} else {
			training_file = training_result_file
			read_pred_training_1 = read_pred_training & all(file.exists(training_file))
			if (!read_pred_training_1) {
				writeLines("Declared:")
				writeLines(paste0("read_pred_training = ", read_pred_training))
				writeLines(paste0("training_result_file = ", paste(training_result_file, collapse=";")))
				stop("Reading existing training result but there is at least one missing training result file. Please check parameter all files available in <training_result_file> or set read_pred_testing=F and training_result_file=NULL")
			}
			writeLines(paste0("Reading all files in training_result_file"))
		}
		
		if (!read_pred_training_1) {

			sumscore = apply(as.data.frame(train_df[,score_names]), 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) train_df = train_df[,-match(names(idx), colnames(train_df))]

			pgs_list_all = colnames(train_df)
			pgs_list_all = pgs_list_all[which(pgs_list_all %in% colnames(all_scores)[2:ncol(all_scores)])]
			pred_acc_train_allPGS_summary = eval_multiple_PRS(train_df, pgs_list_all, covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)
			head(pred_acc_train_allPGS_summary)

			fwrite(pred_acc_train_allPGS_summary, training_file, row.names=F, sep="\t", quote=F)
		} else {
			pred_acc_train_allPGS_summary = NULL
			for (file_i in 1:length(training_file)) {
				writeLines(paste0("Reading training file: ", training_file[file_i]))
				pred_acc_train_allPGS_summary_tmp = fread(training_file[file_i])
				pred_acc_train_allPGS_summary = rbind(pred_acc_train_allPGS_summary, pred_acc_train_allPGS_summary_tmp)
			}
			pred_acc_train_allPGS_summary = pred_acc_train_allPGS_summary[!duplicated(pred_acc_train_allPGS_summary$pgs),]
		}

		pred_acc_train_allPGS_summary = as.data.frame(pred_acc_train_allPGS_summary)

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary
		pred_acc_train_trait_summary = pred_acc_train_trait_summary[order(as.numeric(pred_acc_train_trait_summary$pval), decreasing=F),]
		head(pred_acc_train_trait_summary)

		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary1[order(as.numeric(pred_acc_train_allPGS_summary1$R2), decreasing=T),]
		bestPRS = pred_acc_train_allPGS_summary1[1,1]
		writeLines(paste0("The best single trait-specific score in the training set is ", bestPRS))
		bestPRS_acc = eval_single_PRS(test_df, pheno = "trait", prs_name=bestPRS, covar_list=covar_list, liabilityR2, alpha=0.05, isbinary=isbinary)

		fwrite(bestPRS_acc, paste0(out, "_best_acc.txt"), row.names=F, sep="\t", quote=F)

		################################ testing #################################

		writeLines("--- Evaluating PRS in testing set ---")
		testing_file = paste0(out, "_test_allPRS.txt")
		read_pred_testing_1 = (read_pred_testing & file.exists(testing_file))
		if (!read_pred_testing_1) {
			sumscore = apply(test_df[,score_names], 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) test_df = test_df[,-match(names(idx), colnames(test_df))]

			pred_acc_test_trait = eval_multiple_PRS(test_df, pgs_list,  covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)

			pred_acc_test_trait_summary = pred_acc_test_trait
			pred_acc_test_trait_summary = pred_acc_test_trait_summary[order(as.numeric(pred_acc_test_trait_summary$pval), decreasing=F),]
			print(head(pred_acc_test_trait_summary))

			fwrite(pred_acc_test_trait_summary, testing_file, row.names=F, sep="\t", quote=F)
		} else {
			writeLines(paste0("Reading testing file: ", testing_file))
			pred_acc_test_trait_summary = fread(testing_file)
		}

		###########################################################################
		
		if (isbinary) {

			x_train = (train_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_train = as.vector(train_df$trait)
			train_data = data.frame(x_train,trait=y_train)

			x_test = (test_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_test = as.vector(test_df$trait)
			test_data = data.frame(x_test,trait=y_test)

			train_tmp = train_data[,c("trait", c(bestPRS, covar_list))]
			train_tmp$trait = as.factor(train_tmp$trait)
			
			formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, collapse="+")))
			formula_bestPRS = as.formula(paste0("trait ~ scale(", bestPRS, ")+", paste0(covar_list, collapse="+")))
			
			ctrl = trainControl(
				method = "repeatedcv",
				allowParallel = TRUE,
				number = nfold_cv,
				returnData = FALSE,
				trim = T,
				verboseIter = T)

			cl = makePSOCKcluster(ncores)
			cl = makeCluster(ncores)
			registerDoParallel(cl)

			set.seed(123)
			model_prsmix_null = train(
				formula_null, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)
			
			set.seed(123)
			model_prsmix_bestPRS = train(
				formula_bestPRS, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)

			stopCluster(cl)

			test_data1 = test_data
			
			# --- Null model AUC + ROC ---
      test_pred = predict(model_prsmix_null, test_data1, type = "prob")[,2]
      roc_null <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_null)
      auc_out = data.frame(method="null_model", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_NULL.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_null, paste0(out, "_ROC_null.rds"))
      
      # --- BestPRS model AUC + ROC ---
      test_pred = predict(model_prsmix_bestPRS, test_data1, type = "prob")[,2]
      roc_bestPRS <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_bestPRS)
      auc_out = data.frame(method="bestPGS", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_BestPGS.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_bestPRS, paste0(out, "_ROC_bestPRS.rds"))
			

		}

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		
		pred_acc_test_trait_summary_out = pred_acc_test_trait_summary

		writeLines("--- Iterating power and p-value parameters --- ")
		for (power_thres in power_thres_list)
			for (pval_thres in pval_thres_list) {
				
				writeLines(paste0("P = ", pval_thres))
				writeLines(paste0("Power = ", power_thres))
				writeLines("PRSmix:")

				topprs = pred_acc_train_trait_summary #%>%
					#filter(pval <= pval_thres & power >= power_thres)
				head(topprs)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))
				
				start_time = Sys.time()
				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix")
					ww_raw = 1
				} else {

					if (!isbinary) {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])					
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]

						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}
						
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)

						res_lm1 = eval_single_PRS(test_df1, pheno = "trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

					} else {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])						
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))
						
						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww_raw = ww

							# print(ww)
							# print(sd_train)
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww[which(!names(ww) %in% covar_list)]
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}

							if (debug) {
								print(ww)
							}
							# print(ww)
							test_data1 = test_data
							test_data1[,topprs] = as.numeric(scale(test_data[,topprs]))

							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmix", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmix.txt"), row.names=F, sep="\t", quote=F)
							
						}

						# print(head(test_data))	
						# print(head(test_df))	
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						
						res_lm1 = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

						
						############## OR ###################
						ff = paste0("trait ~ scale(newprs) + ", paste0(covar_list, collapse="+"))
						model1 = glm(ff, data=test_df1, family="binomial")
						model1s = summary(model1)
						mm = exp(model1s$coefficients[2,1])
						ll = exp(model1s$coefficients[2,1] - 1.97*model1s$coefficients[2,2])
						uu = exp(model1s$coefficients[2,1] + 1.97*model1s$coefficients[2,2])
						pval = format.pval(model1s$coefficients[2,4])
						print(paste0("OR = ", rr(mm), " (", rr(ll), "-", rr(uu), "); P-value=", pval))
						fwrite(data.frame(mm, ll, uu, pval), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_OR_PRSmix.txt"), row.names=F, sep="\t", quote=F)

						####################################

					}
					
					
					print(res_lm1)
					res_lm1_summary = res_lm1
					res_lm1_summary$pgs = "PRSmix"
					pred_acc_test_trait_summary_out = bind_rows(res_lm1, pred_acc_test_trait_summary)
					head(pred_acc_test_trait_summary_out)
					
					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmix.txt"), row.names=F, sep="\t", quote=F)
					
					prs_out = test_df1 %>%
						select(IID, newprs)
					colnames(prs_out) = c("IID", "prsmix")
					
					fwrite(prs_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmix.txt"), row.names=F, sep="\t", quote=F)
					
				}

				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]

				timedf = data.frame(pgs="PRSmix", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmix.txt"), row.names=F, sep="\t", quote=F)		

				############################

				writeLines("PRSmix+:")

				topprs = pred_acc_train_allPGS_summary #%>%
					#filter(pval <= pval_thres & power >= power_thres)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))

				start_time = Sys.time()

				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix+")
					ww_raw = 1
				} else {

					if (!isbinary) {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm

						nonzero_w = names(ww[which(ww!=0)])

					} else {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)
						
						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)
							
							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww==0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]							
							}
							
							test_data1 = test_data
							test_data1[,topprs] = scale(test_data[,topprs])
							
							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							roc_PRSMixPlus <- roc(test_data1$trait, test_pred)
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmixP", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", "newprs", covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm
						
						nonzero_w = names(ww[which(ww!=0)])
						
					}
					
					print(res_lm)
					pred_acc_test_trait_summary_out = bind_rows(res_lm, pred_acc_test_trait_summary_out)
					head(pred_acc_test_trait_summary_out)

					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

					prsmixplus = test_df1 %>% select(IID, newprs)

					fwrite(prsmixplus, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmixPlus.txt"), row.names=F, sep="\t", quote=F)

					fwrite(data.frame(topprs, ww), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)					
					
					if (is_extract_adjSNPeff) {
						mixing_weight_file = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt")
						outfile = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_adjSNPeff_PRSmixPlus.txt")
						extract_adjSNPeff(mixing_weight_file, original_beta_files_list, outfile)
					}
					
					fwrite(data.frame(topprs, ww_raw), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_raw_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
					
					##############################################
				}


				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]
				timedf = data.frame(pgs="PRSmix+", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
				

			}
	}
	writeLines("Finished")

```


```{r delong_test_diabetes}
# Compute CIs
auc_ci_null <- ci.auc(roc_null)
auc_ci_best <- ci.auc(roc_bestPRS)
auc_ci_prsmixplus <- ci.auc(roc_PRSMixPlus)


# Save AUC + CI
auc_out <- data.frame(
  method = c("null_model", "bestPGS", "PRSMix+"),
  auc = c(auc_ci_null[2], auc_ci_best[2], auc_ci_prsmixplus[2]),
  lowerCI = c(auc_ci_null[1], auc_ci_best[1], auc_ci_prsmixplus[1]),
  upperCI = c(auc_ci_null[3], auc_ci_best[3], auc_ci_prsmixplus[3])
)
#fwrite(auc_out, paste0(out, "_auc_summary.txt"), sep = "\t", quote = FALSE)

# 🧠 Compare AUCs using DeLong's test
auc_test1 <- roc.test(roc_null, roc_bestPRS, method = "delong")
auc_test2 <- roc.test(roc_null, roc_PRSMixPlus, method = "delong")

print(auc_test1)
print(auc_test2)
```

```{r diabetes_plot, fig.height = 6, fig.width = 6}
plot(
  roc_null,
  col = "#00A087FF", lwd = 2, legacy.axes = TRUE,
  main = "Diabetes" #,
  #print.auc = TRUE, print.auc.y = 0.4, print.auc.x = 0.6
)

plot(
  roc_bestPRS,
  col = "#4DBBD5FF", lwd = 2, add = TRUE #,
  #print.auc = TRUE, print.auc.y = 0.35, print.auc.x = 0.6
)

plot(
  roc_PRSMixPlus,
  col = "#E64B35FF", lwd = 2, add = TRUE#,
  #print.auc = TRUE, print.auc.y = 0.3, print.auc.x = 0.6
)

# Correct the legend: match length of colors/lwd with items
legend(
  "bottomright",
  legend = c(
    sprintf("Covariates Only (AUC = %.3f)", auc(roc_null)),
    sprintf("Covariates + Best PGS (AUC = %.3f)", auc(roc_bestPRS)),
    sprintf("Covariates + PRSMix+ (AUC = %.3f)", auc(roc_PRSMixPlus)),
    sprintf("BestPRS-Null: p = %.3g", auc_test1$p.value),
    sprintf("PRSMix+ - Null: p = %.3g", auc_test2$p.value)
  ),
  col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", NA, NA),
  lwd = c(2, 2, 2, NA, NA),
  bty = "n",
  text.col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", "black", "black")
)
# Save as a plot object
roc_plot_obj_diabetes <- recordPlot()


```

## Any CMD risk factor

```{r}
pheno_name <- "inc_anycmd"

if(dict[[pheno_name]]=="M"){
  categorical_vars <- categorical_vars[categorical_vars !="prev_anycmd"]
}else{
  categorical_vars <-  categorical_vars[categorical_vars !="prev_anycvd"]
}

covariates <- c(categorical_vars, continuous_vars)

pheno_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_"%&% pheno_name %&% "_phenotype.txt"
covariate_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_"%&% pheno_name %&% "_covariates.txt"
pgs_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/" %&% pheno_name %&% "_file_paths.txt"
score_files_list <-fread(pgs_file, header =F)$V1 
trait_specific_score_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/" %&% pheno_name %&% "_PGS_list.txt"
original_beta_files_list = "/bd-fs-mnt/TenantShare/exthome/home/l578311/janise_share/weights.txt"
covar_list = covariates
cat_covar_list = categorical_vars

allPGS_list = NULL
metascore = NULL
liabilityR2 = T
IID_pheno = "IID"
ncores = 1
is_extract_adjSNPeff = F
train_size_list = NULL
training_result_file = NULL
power_thres_list = c(0.95)
pval_thres_list = 0.1 #(0.05/7)
nfold_cv = 3
read_pred_training = FALSE
read_pred_testing = FALSE
debug = F
out = "/bd-fs-mnt/TenantShare/exthome/home/l578311/working_test_dir/"
isbinary = T

original_beta_files_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/janise_share/weights.txt"
isbinary <- TRUE
allPGS_list = NULL
metascore = NULL 
liabilityR2 = TRUE
train_size_list = NULL
training_result_file = NULL
read_pred_training = FALSE
read_pred_testing = FALSE
debug = F 
```

```{r intro_chunk_anycmd}
basic_data = fread(covariate_file)
basic_data = basic_data[,c(IID_pheno, covar_list)]
if (length(cat_covar_list)>0) {

	writeLines("Generating dummy variables from categorical variables")
	basic_data_list = make_dummy_columns(basic_data, cat_covar_list)

	covar_list = covar_list[-match(cat_covar_list, covar_list)]
	covar_list = c(covar_list, basic_data_list[[2]])

	basic_data = basic_data_list[[1]]
}

writeLines("--- Reading all polygenic risk scores ---")
writeLines("Please make sure the column names for the scores ends with _SUM")

all_scores = NULL

for (score_file_i in 1:length(score_files_list)) {

	print(score_file_i)
	score_file = score_files_list[score_file_i]
	dd = fread(score_file)
	idx = which(endsWith(colnames(dd), "_SUM") & colnames(dd)!="NAMED_ALLELE_DOSAGE_SUM")
	idx2 = which(colnames(dd)=="IID")

	dd_sub = dd[,c(idx2,idx)]
	print(dim(dd_sub))
	if (is.null(all_scores)) {
		all_scores = dd_sub
	} else
		all_scores = merge(all_scores, dd_sub, by="IID")
}

colnames(all_scores)[2:ncol(all_scores)] = substring(colnames(all_scores)[2:ncol(all_scores)], 1, nchar(colnames(all_scores)[2:ncol(all_scores)])-4)

if (!is.null(allPGS_list)) {
	pgs_extract = intersect(colnames(all_scores[2:ncol(all_scores)]), allPGS_list)
	all_scores = all_scores[,c(1,match(pgs_extract, colnames(all_scores)))]
}

score_names = colnames(all_scores)[2:ncol(all_scores)]

pgs_list = NULL
for (ff_i in 1:length(trait_specific_score_file)) {
	writeLines(paste0("Reading: ", trait_specific_score_file[ff_i]))
	pgs_list_tmp = fread(trait_specific_score_file[ff_i], header=F)[,1]
	pgs_list = c(pgs_list, pgs_list_tmp)
}
pgs_list = intersect(pgs_list, colnames(all_scores))

pheno = fread(pheno_file)
idx = which(colnames(pheno) %in% c(IID_pheno, pheno_name))
pheno = pheno[,idx]
colnames(pheno) = c("IID", "trait")

writeLines("--- Merging Phenotype and PRS files ---")

pheno_prs = merge(pheno, all_scores, by="IID")
pheno_prs_cov = merge(pheno_prs, basic_data, by.x="IID", by.y=IID_pheno)
pheno_prs_cov = pheno_prs_cov[which(!is.na(pheno_prs_cov$trait)),]

#######################

irnt = function(x) return(qnorm((rank(x,na.last="keep")-0.5)/sum(!is.na(x))))
rr = function(x,d=3) round(x,d)

#######################

null_train_size_list = F
if (is.null(train_size_list)) {
	writeLines("--- Selecting 80% as training data ---")
	train_size_list = floor(0.8*nrow(pheno_prs_cov))
	null_train_size_list = T
} else {
	writeLines("Using custom data split!")
}

out_save = out
```


```{r original_code_anycmd}
for (train_size in train_size_list) {
  
    ################################ Set up Initial Data #################################
		train_size = train_size_list[1]
		writeLines(paste("--- Using ", train_size, " individuals for training sample ---"))
		if (!null_train_size_list) out = paste0(out_save, "_train.", train_size)
		if (isbinary) fwrite(as.data.frame(table(pheno_prs_cov$trait)), paste0(out, "_case_counts.txt"), row.names=F, sep="\t", quote=F)
		set.seed(1)
		train_idx = sample(1:nrow(pheno_prs_cov), train_size)
		train_df = pheno_prs_cov[train_idx,]
		test_df = pheno_prs_cov[-train_idx,]

		cc = apply(train_df[,covar_list], 2, function(x) length(unique(x)))
		covar_list1 = covar_list[-which(cc<=5)]
		for (i in covar_list1) train_df[i] = as.numeric(scale(train_df[i]))
		for (i in covar_list1) test_df[i] = as.numeric(scale(test_df[i]))

		fwrite(train_df[,c(1,2)], paste0(out, "_train_df.txt"), row.names=F, quote=F, sep="\t")
		fwrite(test_df[,c(1,2)], paste0(out, "_test_df.txt"), row.names=F, quote=F, sep="\t")

		################################ training #################################

		writeLines("--- Evaluating PRS in training set ---")
		
		if (!read_pred_training | is.null(training_result_file)) {
			training_file = paste0(out, "_train_allPRS.txt")
			read_pred_training_1 = (read_pred_training & file.exists(training_file))
		} else {
			training_file = training_result_file
			read_pred_training_1 = read_pred_training & all(file.exists(training_file))
			if (!read_pred_training_1) {
				writeLines("Declared:")
				writeLines(paste0("read_pred_training = ", read_pred_training))
				writeLines(paste0("training_result_file = ", paste(training_result_file, collapse=";")))
				stop("Reading existing training result but there is at least one missing training result file. Please check parameter all files available in <training_result_file> or set read_pred_testing=F and training_result_file=NULL")
			}
			writeLines(paste0("Reading all files in training_result_file"))
		}
		
		if (!read_pred_training_1) {

			sumscore = apply(as.data.frame(train_df[,score_names]), 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) train_df = train_df[,-match(names(idx), colnames(train_df))]

			pgs_list_all = colnames(train_df)
			pgs_list_all = pgs_list_all[which(pgs_list_all %in% colnames(all_scores)[2:ncol(all_scores)])]
			pred_acc_train_allPGS_summary = eval_multiple_PRS(train_df, pgs_list_all, covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)
			head(pred_acc_train_allPGS_summary)

			fwrite(pred_acc_train_allPGS_summary, training_file, row.names=F, sep="\t", quote=F)
		} else {
			pred_acc_train_allPGS_summary = NULL
			for (file_i in 1:length(training_file)) {
				writeLines(paste0("Reading training file: ", training_file[file_i]))
				pred_acc_train_allPGS_summary_tmp = fread(training_file[file_i])
				pred_acc_train_allPGS_summary = rbind(pred_acc_train_allPGS_summary, pred_acc_train_allPGS_summary_tmp)
			}
			pred_acc_train_allPGS_summary = pred_acc_train_allPGS_summary[!duplicated(pred_acc_train_allPGS_summary$pgs),]
		}

		pred_acc_train_allPGS_summary = as.data.frame(pred_acc_train_allPGS_summary)

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary
		pred_acc_train_trait_summary = pred_acc_train_trait_summary[order(as.numeric(pred_acc_train_trait_summary$pval), decreasing=F),]
		head(pred_acc_train_trait_summary)

		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary1[order(as.numeric(pred_acc_train_allPGS_summary1$R2), decreasing=T),]
		bestPRS = pred_acc_train_allPGS_summary1[1,1]
		writeLines(paste0("The best single trait-specific score in the training set is ", bestPRS))
		bestPRS_acc = eval_single_PRS(test_df, pheno = "trait", prs_name=bestPRS, covar_list=covar_list, liabilityR2, alpha=0.05, isbinary=isbinary)

		fwrite(bestPRS_acc, paste0(out, "_best_acc.txt"), row.names=F, sep="\t", quote=F)

		################################ testing #################################

		writeLines("--- Evaluating PRS in testing set ---")
		testing_file = paste0(out, "_test_allPRS.txt")
		read_pred_testing_1 = (read_pred_testing & file.exists(testing_file))
		if (!read_pred_testing_1) {
			sumscore = apply(test_df[,score_names], 2, var)
			idx = which(sumscore==0)
			if (length(idx)>0) test_df = test_df[,-match(names(idx), colnames(test_df))]

			pred_acc_test_trait = eval_multiple_PRS(test_df, pgs_list,  covar_list, liabilityR2, alpha=0.05, isbinary=isbinary, ncores=ncores)

			pred_acc_test_trait_summary = pred_acc_test_trait
			pred_acc_test_trait_summary = pred_acc_test_trait_summary[order(as.numeric(pred_acc_test_trait_summary$pval), decreasing=F),]
			print(head(pred_acc_test_trait_summary))

			fwrite(pred_acc_test_trait_summary, testing_file, row.names=F, sep="\t", quote=F)
		} else {
			writeLines(paste0("Reading testing file: ", testing_file))
			pred_acc_test_trait_summary = fread(testing_file)
		}

		###########################################################################
		
		if (isbinary) {

			x_train = (train_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_train = as.vector(train_df$trait)
			train_data = data.frame(x_train,trait=y_train)

			x_test = (test_df %>% select(all_of(c(bestPRS, covar_list)), -trait))
			y_test = as.vector(test_df$trait)
			test_data = data.frame(x_test,trait=y_test)

			train_tmp = train_data[,c("trait", c(bestPRS, covar_list))]
			train_tmp$trait = as.factor(train_tmp$trait)
			
			formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, collapse="+")))
			formula_bestPRS = as.formula(paste0("trait ~ scale(", bestPRS, ")+", paste0(covar_list, collapse="+")))
			
			ctrl = trainControl(
				method = "repeatedcv",
				allowParallel = TRUE,
				number = nfold_cv,
				returnData = FALSE,
				trim = T,
				verboseIter = T)

			cl = makePSOCKcluster(ncores)
			cl = makeCluster(ncores)
			registerDoParallel(cl)

			set.seed(123)
			model_prsmix_null = train(
				formula_null, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)
			
			set.seed(123)
			model_prsmix_bestPRS = train(
				formula_bestPRS, data = train_tmp, method = "glmnet",
				trControl = ctrl, family = "binomial",
				tuneLength = 50, verbose=T
			)

			stopCluster(cl)

			test_data1 = test_data
			
			# --- Null model AUC + ROC ---
      test_pred = predict(model_prsmix_null, test_data1, type = "prob")[,2]
      roc_null <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_null)
      auc_out = data.frame(method="null_model", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_NULL.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_null, paste0(out, "_ROC_null.rds"))
      
      # --- BestPRS model AUC + ROC ---
      test_pred = predict(model_prsmix_bestPRS, test_data1, type = "prob")[,2]
      roc_bestPRS <- roc(test_data1$trait, test_pred)
      auc_ci = ci.auc(roc_bestPRS)
      auc_out = data.frame(method="bestPGS", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
      fwrite(auc_out, paste0(out, "_auc_BestPGS.txt"), row.names=F, sep="\t", quote=F)
      saveRDS(roc_bestPRS, paste0(out, "_ROC_bestPRS.rds"))
			

		}

		pred_acc_train_trait_summary = pred_acc_train_allPGS_summary %>%
			filter(pgs %in% pgs_list)
		
		pred_acc_test_trait_summary_out = pred_acc_test_trait_summary

		writeLines("--- Iterating power and p-value parameters --- ")
		for (power_thres in power_thres_list)
			for (pval_thres in pval_thres_list) {
				
				writeLines(paste0("P = ", pval_thres))
				writeLines(paste0("Power = ", power_thres))
				writeLines("PRSmix:")

				topprs = pred_acc_train_trait_summary %>%
					filter(pval <= pval_thres & power >= power_thres)
				head(topprs)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))
				
				start_time = Sys.time()
				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix")
					ww_raw = 1
				} else {

					if (!isbinary) {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])					
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]

						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}
						
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)

						res_lm1 = eval_single_PRS(test_df1, pheno = "trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

					} else {
						
						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])						
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))
						
						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = ww_raw = c(1)
							names(ww) = names(ww_raw) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww_raw = ww

							# print(ww)
							# print(sd_train)
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww[which(!names(ww) %in% covar_list)]
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}

							if (debug) {
								print(ww)
							}
							# print(ww)
							test_data1 = test_data
							test_data1[,topprs] = as.numeric(scale(test_data[,topprs]))

							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmix", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmix.txt"), row.names=F, sep="\t", quote=F)
							
						}

						# print(head(test_data))	
						# print(head(test_df))	
						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						
						res_lm1 = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						
						res_lm1$pgs = "PRSmix"
						res_lm1

						
						############## OR ###################
						ff = paste0("trait ~ scale(newprs) + ", paste0(covar_list, collapse="+"))
						model1 = glm(ff, data=test_df1, family="binomial")
						model1s = summary(model1)
						mm = exp(model1s$coefficients[2,1])
						ll = exp(model1s$coefficients[2,1] - 1.97*model1s$coefficients[2,2])
						uu = exp(model1s$coefficients[2,1] + 1.97*model1s$coefficients[2,2])
						pval = format.pval(model1s$coefficients[2,4])
						print(paste0("OR = ", rr(mm), " (", rr(ll), "-", rr(uu), "); P-value=", pval))
						fwrite(data.frame(mm, ll, uu, pval), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_OR_PRSmix.txt"), row.names=F, sep="\t", quote=F)

						####################################

					}
					
					
					print(res_lm1)
					res_lm1_summary = res_lm1
					res_lm1_summary$pgs = "PRSmix"
					pred_acc_test_trait_summary_out = bind_rows(res_lm1, pred_acc_test_trait_summary)
					head(pred_acc_test_trait_summary_out)
					
					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmix.txt"), row.names=F, sep="\t", quote=F)
					
					prs_out = test_df1 %>%
						select(IID, newprs)
					colnames(prs_out) = c("IID", "prsmix")
					
					fwrite(prs_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmix.txt"), row.names=F, sep="\t", quote=F)
					
				}

				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]

				timedf = data.frame(pgs="PRSmix", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmix.txt"), row.names=F, sep="\t", quote=F)		

				############################

				writeLines("PRSmix+:")

				topprs = pred_acc_train_allPGS_summary %>%
					filter(pval <= pval_thres & power >= power_thres)
				topprs = topprs$pgs
				topprs = intersect(topprs, colnames(train_df))
				print(length(topprs))

				start_time = Sys.time()

				if (length(topprs) == 0) {
					print("No high power trait-specific PRS for PRSmix+")
					ww_raw = 1
				} else {

					if (!isbinary) {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)

						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {

							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)

							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)

							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl,
								tuneLength = 50, verbose=T
							)

							stopCluster(cl)

							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww == 0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]
							}
						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", prs_name="newprs", covar_list=covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm

						nonzero_w = names(ww[which(ww!=0)])

					} else {

						x_train = (train_df %>% select(all_of(c(topprs, covar_list)), -trait))
						if (length(topprs) > 1)	sd_train = apply(as.data.frame(x_train[,topprs]), 2, sd, na.rm=T)
						x_train[,topprs] = scale(x_train[,topprs])
						y_train = as.vector(train_df$trait)
						train_data = data.frame(x_train,trait=y_train)

						x_test = (test_df %>% select(all_of(c(topprs, covar_list)), -trait))
						y_test = as.vector(test_df$trait)
						test_data = data.frame(x_test,trait=y_test)
						
						formula = as.formula(paste0("trait ~ ", paste0(topprs, collapse="+"), "+", paste0(covar_list, collapse="+")))

						train_tmp = train_data[,c("trait", topprs, covar_list)]
						train_tmp$trait = as.factor(train_tmp$trait)
						writeLines("Header of training data")
						print(head(train_tmp))

						if (length(topprs) == 1) {
							ww = c(1)
							names(ww) = topprs
						} else {
							
							ctrl = trainControl(
								method = "repeatedcv",
								allowParallel = TRUE,
								number = nfold_cv,
								verboseIter = T)
							
							cl = makePSOCKcluster(ncores)
							registerDoParallel(cl)
							
							set.seed(123)
							model_prsmix = train(
								formula, data = train_tmp, method = "glmnet",
								trControl = ctrl, family = "binomial",
								tuneLength = 50, verbose=T
							)
							
							stopCluster(cl)
							
							model_prsmix$bestTune
							ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[,1][-1]
							ww = ww[which(!names(ww) %in% covar_list)]
							ww_raw = ww							
							
							if (all(ww==0)) {
								writeLines("No weight for PRS")
								ww = c(1)
								names(ww) = bestPRS_acc$pgs
								topprs = bestPRS_acc$pgs
							} else {
								ww = ww / sd_train[match(names(ww), names(sd_train))]							
							}
							
							test_data1 = test_data
							test_data1[,topprs] = scale(test_data[,topprs])
							
							test_pred = predict(model_prsmix, test_data1, type = "prob")[,2]
							roc_PRSMixPlus <- roc(test_data1$trait, test_pred)
							auc_ci = ci.auc(test_data1$trait, test_pred)
							auc_out = data.frame(method="prsmixP", auc=auc_ci[2], lowerCI=auc_ci[1], upperCI=auc_ci[3])
							fwrite(auc_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_auc_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

						}

						test_df1 = cbind(test_data, IID=test_df$IID)
						test_df1$newprs = as.matrix(test_df1[,topprs]) %*% as.vector(ww)
						res_lm = eval_single_PRS(test_df1, pheno="trait", "newprs", covar_list, liabilityR2 = liabilityR2, alpha=pval_thres, isbinary=isbinary)
						res_lm$pgs = "PRSmix+"
						res_lm
						
						nonzero_w = names(ww[which(ww!=0)])
						
					}
					
					print(res_lm)
					pred_acc_test_trait_summary_out = bind_rows(res_lm, pred_acc_test_trait_summary_out)
					head(pred_acc_test_trait_summary_out)

					fwrite(pred_acc_test_trait_summary_out, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_test_summary_traitPRS_withPRSmixPlus.txt"), row.names=F, sep="\t", quote=F)

					prsmixplus = test_df1 %>% select(IID, newprs)

					fwrite(prsmixplus, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_prsmixPlus.txt"), row.names=F, sep="\t", quote=F)

					fwrite(data.frame(topprs, ww), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)					
					
					if (is_extract_adjSNPeff) {
						mixing_weight_file = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt")
						outfile = paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_adjSNPeff_PRSmixPlus.txt")
						extract_adjSNPeff(mixing_weight_file, original_beta_files_list, outfile)
					}
					
					fwrite(data.frame(topprs, ww_raw), paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_weight_raw_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
					
					##############################################
				}


				end_time = Sys.time()
				timerunning = difftime(end_time, start_time, units = "secs")[[1]]
				timedf = data.frame(pgs="PRSmix+", npgs=length(ww_raw), time=timerunning)
				print(timedf)

				fwrite(timedf, paste0(out, "_power.", power_thres, "_pthres.", pval_thres, "_time_PRSmixPlus.txt"), row.names=F, sep="\t", quote=F)
				

			}
	}
	writeLines("Finished")

```


```{r delong_test_anycmd}
# Compute CIs
auc_ci_null <- ci.auc(roc_null)
auc_ci_best <- ci.auc(roc_bestPRS)
auc_ci_prsmixplus <- ci.auc(roc_PRSMixPlus)


# Save AUC + CI
auc_out <- data.frame(
  method = c("null_model", "bestPGS", "PRSMix+"),
  auc = c(auc_ci_null[2], auc_ci_best[2], auc_ci_prsmixplus[2]),
  lowerCI = c(auc_ci_null[1], auc_ci_best[1], auc_ci_prsmixplus[1]),
  upperCI = c(auc_ci_null[3], auc_ci_best[3], auc_ci_prsmixplus[3])
)
#fwrite(auc_out, paste0(out, "_auc_summary.txt"), sep = "\t", quote = FALSE)

# 🧠 Compare AUCs using DeLong's test
auc_test1 <- roc.test(roc_null, roc_bestPRS, method = "delong")
auc_test2 <- roc.test(roc_null, roc_PRSMixPlus, method = "delong")

print(auc_test1)
print(auc_test2)
```

```{r anycmd_plot, fig.height = 6, fig.width = 6}
plot(
  roc_null,
  col = "#00A087FF", lwd = 2, legacy.axes = TRUE,
  main = "Any CMD" #,
  #print.auc = TRUE, print.auc.y = 0.4, print.auc.x = 0.6
)

plot(
  roc_bestPRS,
  col = "#4DBBD5FF", lwd = 2, add = TRUE #,
  #print.auc = TRUE, print.auc.y = 0.35, print.auc.x = 0.6
)

plot(
  roc_PRSMixPlus,
  col = "#E64B35FF", lwd = 2, add = TRUE#,
  #print.auc = TRUE, print.auc.y = 0.3, print.auc.x = 0.6
)

# Correct the legend: match length of colors/lwd with items
legend(
  "bottomright",
  legend = c(
    sprintf("Covariates Only (AUC = %.3f)", auc(roc_null)),
    sprintf("Covariates + Best PGS (AUC = %.3f)", auc(roc_bestPRS)),
    sprintf("Covariates + PRSMix+ (AUC = %.3f)", auc(roc_PRSMixPlus)),
    sprintf("BestPRS-Null: p = %.3g", auc_test1$p.value),
    sprintf("PRSMix+ - Null: p = %.3g", auc_test2$p.value)
  ),
  col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", NA, NA),
  lwd = c(2, 2, 2, NA, NA),
  bty = "n",
  text.col = c("#00A087FF", "#4DBBD5FF", "#E64B35FF", "black", "black")
)
# Save as a plot object
roc_plot_obj_anycmd <- recordPlot()


```

# Final Plots


```{r all_plots, fig.height= 12, fig.width=18}
library(ggplotify)
library(cowplot)

# Convert recordedplot → ggplot objects

# Convert each recordedplot object safely to ggplot
p1 <- as.ggplot(~replayPlot(roc_plot_obj_ischemic))
p2 <- as.ggplot(~replayPlot(roc_plot_obj_htn))
p3 <- as.ggplot(~replayPlot(roc_plot_obj_dyslipid))
p4 <- as.ggplot(~replayPlot(roc_plot_obj_diabetes))
p5 <- as.ggplot(~replayPlot(roc_plot_obj_anycmd))
# Combine them with plot_grid()
plot_grid(p1, p2, p3, p4, p5, ncol = 3)


```

