---
title: "PGS_Processing Notes"
output: html_document
date: "2025-04-23"
---

```{r, echo = FALSE}
library(dplyr)
library(tidyr)
library(data.table)
library(readxl)
library(ggplot2)
library(ggsci)

"%&%"=function(a,b) paste(a,b,sep="")

dir <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/"
```

# Introduction

This document serves as my notes for performing the updated PGS analyses on KPNC ACI. Previously, I set these scripts up to run on an HPC Cluster with a SLURM or torque scheduler, but that is not the computing infrastructure I am using at KPNC. Instead, I will be updating the code to run on simpler sessions of Rstudio and Linux.

Additionally, I will be picking up with the script `03_process_phenotype.R`. The scripts before this entail the processing of genotypes, PGS calculation, and PGS processing.  

# Phenotype Processing

The goal of `03_process_phenotype.R` is to update the BMI, income, and education coding. After this, all the phenotypic information is joined to the genotype information in the PGS. It will then write one large phenotype file, individual phenotype files, and then one large phenotype file with the covariates. It then does all of these for the ancestry-specific information.

It looks like Ally already joined the PGS to the phenotype data in the files located at `/bd-fs-mnt/TenantShare/exthome/home/l578311/cvd/pgs_revised/data/final_cohorts/`.

# Pull PGS Outcomes

For the fourth script, I had to have Janise copy some additional files and folders from CCR to KPNC. This script updates some phenotype names and formats the PGS into PGS number rather than the trait name.

# Run PRSMix

I realized in this previous step that I was only using the phenotypic (STUDYIDs) and not the genotype information. I shared a Genotype map file with Janise, and changed the names in the phenotype to be the genotypic ID that is used for all the PGS information. 

From here, I re-ran step `03_KPNC_process_phenotype.R` with a couple lines to update this. From there, I re-ran the `04` script. I then make `04_KPNC_call_ancestry_PRSMix.sh` which runs the analyses for all 5 groups in series. 


The code appears to run without syntax or file errors, but I am getting the error:

```
Aggregating results
Something is wrong; all Accuracy metric vaues are missing:
```

Then all the values are NA or NaN. I discovered this because there are "2"s and "3"s coded in the phenotypes. When I revisited the data dictionary, Ally put these for when someone died before developing the phenotype (2) or had a prevalent phenotype at the time of diagnosis (3).


## Troubleshooting PRSMix

When I look through the log files, there are different errors that come up. For example, in `ALL_inc_dyslipid.log`, the error is:

```
Error in na.fail.default(list(trait = c(2L, 2L, 1L, 1L, 1L, 2L, 2L, 2L,  : 
  missing values in object
Calls: combine_PRS ... model.frame.default -> <Anonymous> -> na.fail.default
Execution halted
```

When I check through these, I find that the lines 302-313 of the `combine_PRS` script where where the `train()` function is used may be causing the error.

Additionally, I find that in `EUR_inc_anycmd.log`, there is the different error:

```
Error in roc.default(response, predictor, ci = FALSE, ...) : 
  Response and predictor must be vectors of the same length.
Calls: combine_PRS -> ci.auc -> ci.auc.default -> roc.default
In addition: Warning message:
In for (i in 1:length(args)) { :
  closing unused connection 4 (<-localhost:11988)
Execution halted
```

This may be from an error at line 320 of the `combine_PRS()` script.



Let's start with the first error:


```{r, eval = FALSE}
covariate_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/ALL__covariates.txt"
basic_data = fread(covariate_file)

train_df <- fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_results/ALL_inc_dyslipid_train_df.txt", header = T)

categorical_vars<- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

IID_pheno <- c("IID")

covar_list <- c(categorical_vars, continuous_vars)

basic_data = fread(covariate_file)
basic_data = basic_data[, c(IID_pheno, covar_list)]

basic_data <- basic_data %>% filter(!is.na(bmi_new))

train_tmp <- merge(train_df, basic_data, by="IID")

train_tmp$trait = as.factor(train_tmp$trait)
formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, 
  collapse = "+")))

nfold_cv = 3
ncores=1
ctrl = trainControl(method = "repeatedcv", allowParallel = TRUE, 
  number = nfold_cv, returnData = FALSE, trim = T, 
  verboseIter = T)
cl = makePSOCKcluster(ncores)
cl = makeCluster(ncores)
registerDoParallel(cl)
set.seed(123)
model_prsmix_null = train(formula_null, data = train_tmp, 
  method = "glmnet", trControl = ctrl, family = "binomial", 
  tuneLength = 50, verbose = T)


formula_null = as.formula(paste0("trait ~", paste0(covar_list, collapse = "+")))

model_prsmix_null = train(
  formula_null, data = train_tmp, method = "glmnet",
  trControl = ctrl, family = "binomial",
  tuneLength = 50, verbose = T
)

```

This was resolved by removing the `NA` from `bmi_new`. After resolving this error. The other error went away... But, a new one showed up:

In all the `inc_anycmd`, there is the issue:

```
--- Reading covariate data ---
Generating dummy variables from categorical variables
Error in top_vals$vals : $ operator is invalid for atomic vectors
Calls: combine_PRS ... make_dummy_columns -> dummy_cols -> <Anonymous> -> stri_order
In addition: Warning messages:
1: In max(vals$Freq) : no non-missing arguments to max; returning -Inf
2: In max(vals$Freq) : no non-missing arguments to max; returning -Inf
Execution halted
```


```{r, echo = FALSE, eval = FALSE}
#COVARIATE
categorical_vars<- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

covariates <- c(categorical_vars, continuous_vars)

scores_file_vector <-fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_anycmd_file_paths.txt", header =F)$V1 

pgs_list <- fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_anycmd_PGS_list.txt", header = F)
num_pgs <- nrow(pgs_list)

#print("Processing the phenotype: " %&% args$pheno_name %&%" ; Number of PGS Corresponding to it: " %&% num_pgs)

combine_PRS(
  pheno_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/ALL_inc_anycmd_phenotype.txt" ,
  covariate_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/ALL_inc_anycmd_covariates.txt" ,
  score_files_list = scores_file_vector ,
  trait_specific_score_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_anycmd_file_paths.txt" ,
  pheno_name = "inc_anycmd" ,
  isbinary = TRUE , 
  out = "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/test", 
  liabilityR2 = TRUE ,
  IID_pheno =  "IID" ,
  covar_list = covariates,
  cat_covar_list = categorical_vars,
  is_extract_adjSNPeff = FALSE,
  train_size_list = NULL,
  #power_thres_list = 0.95,
  pval_thres_list = 0.05/7,
  read_pred_training = FALSE,
  read_pred_testing = FALSE,
  original_beta_files_list = "/bd-fs-mnt/TenantShare/exthome/home/l578311/janise_share/weights.txt"
)

covariate_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/ALL_inc_anycmd_covariates.txt"
IID_pheno="IID"
covar_list = covariates
cat_covar_list = categorical_vars
options(datatable.fread.datatable = FALSE)
#writeLines("--- Reading covariate data ---")
basic_data = fread(covariate_file)
basic_data = basic_data[, c(IID_pheno, covar_list)]
if (length(cat_covar_list) > 0) {
  writeLines("Generating dummy variables from categorical variables")
  basic_data_list = make_dummy_columns(basic_data, cat_covar_list)
  covar_list = covar_list[-match(cat_covar_list, covar_list)]
  covar_list = c(covar_list, basic_data_list[[2]])
  basic_data = basic_data_list[[1]]
}

df = basic_data
dummy_colnames = NULL
for (cat_covar_i in 1:length(cat_covar_list)) {
  print(cat_covar_i)
  cat_covar = cat_covar_list[cat_covar_i]
  df = df[which(!is.na(df[, cat_covar])), ]
  df = dummy_cols(df, select_columns = cat_covar, remove_selected_columns = T, 
    remove_most_frequent_dummy = T)
  dummy_colnames_tmp = colnames(df)[which(startsWith(colnames(df), 
    cat_covar))]
  dummy_colnames = c(dummy_colnames, dummy_colnames_tmp)
}
return(list(df = df, dummy_names = dummy_colnames))
```

The issue above was addressed by fixing the phenotype file processing step. The prev_anycmd was being used for the inc_anycmd phenotype. This created issues because these individuals had to have no previous history.

# PRSMix Results

## Compare Accuracy
```{r}
pgs_dir<-"/bd-fs-mnt/TenantShare/exthome/home/l578311/"
file_directory <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_results"
file_list <- list.files(file_directory, pattern = "*_summary_traitPRS_withPRSmixPlus.txt", full.names = TRUE)
pheno_list <- c()

for (file in file_list) {
  phenotype <- sub(".*/(.*?)_power.0.95_pthres.0.00714285714285714_test_summary_traitPRS_withPRSmixPlus.txt", "\\1", file)
  pheno_list <- c(pheno_list, phenotype)
}

auc_pgs <- data.table()
auc_null <- data.table()
for (pheno in pheno_list){
  #AUC
  a_auc_pgs <- fread(pgs_dir %&% "prsmix_results/" %&% pheno %&% "_auc_BestPGS.txt", header = T)
  a_auc_pgs$pheno <- pheno
  a_auc_null <- fread(pgs_dir %&% "prsmix_results/" %&% pheno %&% "_auc_NULL.txt", header = T)
  a_auc_null$pheno <- pheno
  
  auc_pgs <- bind_rows(auc_pgs, a_auc_pgs)
  auc_null <- bind_rows(auc_null, a_auc_null)
  
  # ACCURACY
  
  
}
auc_comp <- bind_rows(auc_pgs, auc_null)
```

```{r}
ggplot(auc_comp, aes(x = auc, y = pheno, fill = method)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +  # Horizontal bars with borders
  geom_errorbar(aes(xmin = lowerCI, xmax = upperCI), 
                width = 0.3,  # Width of the error bar ends
                position = position_dodge(width = 0.9)) +  # Dodge to align with bars
  geom_text(aes(label = round(auc, 3), x = auc - .1),  # Add labels slightly to the right
        position = position_dodge(width = 0.9),    # Align with bars
        size = 4,                                  # Adjust text size
        hjust = 1) +
  labs(
    x = "AUC",
    y = "Phenotype",
    title = "AUC for Phenotypes by Method"
  ) +
  theme_bw(14) +  # Clean theme
  theme(
    axis.text.y = element_text(size = 10),  # Adjust y-axis text size
    axis.title = element_text(size = 12),  # Adjust axis title size
    legend.position = "right"  # Place legend on the right
  ) +
  scale_fill_npg()
```

### Ancestry Specific
```{r}
pgs_dir <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/"
file_directory <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_results"
file_list <- list.files(file_directory, pattern = "*_summary_traitPRS_withPRSmixPlus.txt", full.names = TRUE)
pheno_list <- c("inc_diabetes" ,"inc_dyslipid" ,"inc_ischemic", "inc_anycmd" , "inc_hyperten" )
#pheno_list <- c("inc_diabetes" ,"inc_dyslipid" ,"inc_ischemic" , "inc_hyperten" )


ANCESTRIES <- c("ALL",  "EUR", "AFR", "EAS", "AMR")

auc_pgs <- data.table()
auc_null <- data.table()
for(ancestry in ANCESTRIES){
  if(ancestry =="ALL"){
        for (pheno in pheno_list){
        #AUC
        a_auc_pgs <- fread(pgs_dir %&% "prsmix_results/" %&% ancestry %&% "_" %&% pheno %&% "_auc_BestPGS.txt", header = T)
        a_auc_pgs$pheno <- pheno
        a_auc_pgs$anc <- ancestry
        a_auc_null <- fread(pgs_dir %&% "prsmix_results/" %&% ancestry %&% "_" %&% pheno %&% "_auc_NULL.txt", header = T)
        a_auc_null$anc <- ancestry
        a_auc_null$pheno <- pheno
        
        auc_pgs <- bind_rows(auc_pgs, a_auc_pgs)
        auc_null <- bind_rows(auc_null, a_auc_null)
      }
    }else{
    for (pheno in pheno_list){
      #AUC
      a_auc_pgs <- fread(pgs_dir %&% "prsmix_results/" %&% ancestry %&% "_" %&% pheno %&% "_auc_BestPGS.txt", header = T)
      a_auc_pgs$pheno <- pheno
      a_auc_pgs$anc <- ancestry
      a_auc_null <- fread(pgs_dir %&% "prsmix_results/" %&% ancestry %&% "_" %&%  pheno %&% "_auc_NULL.txt", header = T)
      a_auc_null$anc <- ancestry
      a_auc_null$pheno <- pheno
      
      auc_pgs <- bind_rows(auc_pgs, a_auc_pgs)
      auc_null <- bind_rows(auc_null, a_auc_null)
    
    # ACCURACY
    }
  }
}


auc_comp <- bind_rows(auc_pgs, auc_null)
auc_comp$anc<-factor(auc_comp$anc, levels = c("ALL", "AFR", "AMR", "EAS", "EUR"))
```


```{r, fig.height= 12}
ggplot(auc_comp, aes(x = auc, y = pheno, fill = anc)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", alpha =0.75) +  # Horizontal bars with borders
  geom_errorbar(aes(xmin = lowerCI, xmax = upperCI), 
                width = 0.3,  # Width of the error bar ends
                position = position_dodge(width = 0.9)) +  # Dodge to align with bars
  geom_text(aes(label = round(auc, 3), x = 0.05),  # Add labels slightly to the right
        position = position_dodge(width = 0.9),    # Align with bars
        size = 4,                                  # Adjust text size
        hjust = 1) +
  labs(
    x = "AUC",
    y = "Phenotype",
    title = "AUC for Phenotypes by Method"
  ) +
  theme_bw(14) +  # Clean theme
  theme(
    axis.text.y = element_text(size = 10),  # Adjust y-axis text size
    axis.title = element_text(size = 12),  # Adjust axis title size
    legend.position = "right"  # Place legend on the right
  ) +
  scale_fill_npg() + facet_wrap(~method, nrow=2)
```

# Pull difference in performance
```{r}

setDT(auc_comp)

# Check if there are both `bestPGS` and `null_model` methods for each group
if (all(c("bestPGS", "null_model") %in% unique(auc_comp$method))) {
  # Calculate the difference in AUC between bestPGS and null_model
  auc_differences <- auc_comp[
    , .(auc_diff = auc[method == "bestPGS"] - auc[method == "null_model"]),
    by = .(pheno, anc)
  ]
  
  # Print the results
  print(auc_differences)
} else {
  print("Both 'bestPGS' and 'null_model' must be present in the data for each group.")
}


ggplot(auc_differences, aes(x = auc_diff, y = pheno, fill = anc)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", alpha =0.75) +  # Horizontal bars with borders
  labs(
    x = "Difference in AUC between PGS and covariate-only model",
    y = "Phenotype",
    title = "AUC Difference for Phenotypes"
  ) +
  theme_bw(14) +  # Clean theme
  theme(
    axis.text.y = element_text(size = 10),  # Adjust y-axis text size
    axis.title = element_text(size = 12),  # Adjust axis title size
    legend.position = "right"  # Place legend on the right
  ) +
  scale_fill_npg()
```


```{r}
file_directory <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_results"
file_list <- list.files(file_directory, pattern = "*_summary_traitPRS_withPRSmix.txt", full.names = TRUE)

prsmixplus_summary <- data.table()
for (file in file_list){
  phenotype <- sub(".*/(.*?)_power.0.95_pthres.0.00714285714285714_test_summary_traitPRS_withPRSmix.txt", "\\1", file)
  start <- fread(file)
  print("Processing: " %&% file)
  start$phenotype <- phenotype
  prsmixplus_summary <- bind_rows(prsmixplus_summary, start)
}

```

```{r, fig.width = 6, fig.height = 10}
ggplot(prsmixplus_summary, aes(x = R2, y = pgs, fill = phenotype)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +  # Horizontal bars with borders
  geom_errorbar(aes(xmin = lowerCI, xmax = upperCI), 
                width = 0.3,  # Width of the error bar ends
                position = position_dodge(width = 0.9)) +  # Dodge to align with bars
  labs(
    x = "Accuracy (R2)",
    y = "PGS",
    title = "R2 for Phenotypes by Method"
  ) +
  theme_bw(14) +  # Clean theme
  theme(
    axis.text.y = element_text(size = 10),  # Adjust y-axis text size
    axis.title = element_text(size = 12),  # Adjust axis title size
    legend.position = "right"  # Place legend on the right
  ) +
  geom_text(aes(label = round(R2, 3), x = R2 + 0.02),  # Add labels slightly to the right
          position = position_dodge(width = 0.9),    # Align with bars
          size = 4,                                  # Adjust text size
          hjust = 1, vjust=1.3) +
  scale_fill_npg() + facet_wrap(~phenotype, scales = "free_y", nrow = 5) +guides(fill="none")
```

1) Cox Model
2) Follow Up to determine best PGS
3) Differences in Performance by Ancestry
4) Comparison to covariates alone.
5) Stratify by recipient of disease.


# Stratification 


## Combined

```{r split_string}
library(stringr)
split_string_parts <- function(x) {
  parts <- str_split(x, "_", simplify = T)
  n <- ncol(parts)
  before_second<- str_c(parts[1:2], collapse = "_")
  between<- str_c(parts[3:(n-1)], collapse = "_")
  after_last<- parts[n]
  return(list(before_second = before_second,
              between= between,
              after_last = after_last))
}
```

```{r}
pgs_dir<-"/bd-fs-mnt/TenantShare/exthome/home/l578311/"

treatment<- c("rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")
pheno_list <- c()
file_directory <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_results/treatment_strat/"
file_list <- list.files(file_directory, pattern = "*auc_BestPGS.txt", full.names = TRUE)
for (file in file_list) {
  phenotype <- sub(".*/ALL_(.*?)_auc_BestPGS.txt", "\\1", file)
  pheno_list <- c(pheno_list, phenotype)
}

# Auc Values
auc_pgs <- data.table()
auc_null <- data.table()
or_pgs <-data.table()
or_null <- data.table()

pgs_acc <- data.table()
for (pheno in pheno_list){
  # ACCURACY
  best_pgs_acc <- fread(file_directory %&% "ALL_" %&% pheno %&% "_best_acc.txt", header = T)
  top_pgs <- best_pgs_acc$pgs
  best_pgs_acc$pheno <- pheno

  pgs_acc <- bind_rows(pgs_acc, best_pgs_acc)
  #AUC
  a_auc_pgs <- fread(pgs_dir %&% "prsmix_results/treatment_strat/ALL_" %&% pheno %&% "_auc_BestPGS.txt", header = T)
  a_auc_pgs$pheno <- pheno
  a_auc_pgs$pgs <- top_pgs
  a_auc_null <- fread(pgs_dir %&% "prsmix_results/treatment_strat/ALL_" %&% pheno %&% "_auc_NULL.txt", header = T)
  a_auc_null$pheno <- pheno
  
  auc_pgs <- bind_rows(auc_pgs, a_auc_pgs)
  auc_null <- bind_rows(auc_null, a_auc_null)
  
  #OR
  a_or_pgs <- fread(file_directory %&% "ALL_" %&% pheno %&% "_OR_BestPGS.txt", header = T) %>%
    mutate(mm = as.numeric(mm))
  a_or_pgs$pheno <- pheno
  a_or_pgs$pgs <- top_pgs
  
  or_pgs <- bind_rows(or_pgs, a_or_pgs)


}
auc_comp <- bind_rows(auc_pgs, auc_null)
auc_comp_sep <- auc_comp %>%
  rowwise() %>%
  mutate(split= list(split_string_parts(pheno))) %>%
  unnest_wider(split, names_sep = "_") %>%
  rename(outcome = split_before_second,
         treatment = split_between,
         received_tx = split_after_last) %>%
  ungroup()

or_comb<- or_pgs %>%
  rowwise() %>%
  mutate(split= list(split_string_parts(pheno))) %>%
  unnest_wider(split, names_sep = "_") %>%
  rename(outcome = split_before_second,
         treatment = split_between,
         received_tx = split_after_last) %>%
  ungroup()

acc_comb <- pgs_acc %>%
  rowwise() %>%
  mutate(split= list(split_string_parts(pheno))) %>%
  unnest_wider(split, names_sep = "_") %>%
  rename(outcome = split_before_second,
         treatment = split_between,
         received_tx = split_after_last) %>%
  ungroup()








  
```

```{r, fig.height = 20}
ggplot(auc_comp_sep, aes(x = auc, y = outcome, fill = interaction(method, received_tx))) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +  # Horizontal bars with borders
  geom_errorbar(aes(xmin = lowerCI, xmax = upperCI), 
                width = 0.3,  # Width of the error bar ends
                position = position_dodge(width = 0.9)) +  # Dodge to align with bars
  geom_text(aes(label = round(auc, 3), x = auc - .1),  # Add labels slightly to the right
        position = position_dodge(width = 0.9),    # Align with bars
        size = 4,                                  # Adjust text size
        hjust = 1) +
  labs(
    x = "AUC",
    y = "Phenotype",
    title = "AUC for Phenotypes by Method"
  ) +
  theme_bw(14) +  # Clean theme
  theme(
    axis.text.y = element_text(size = 10),  # Adjust y-axis text size
    axis.title = element_text(size = 12),  # Adjust axis title size
    legend.position = "right"  # Place legend on the right
  ) +
  scale_fill_npg() + facet_wrap(~treatment, ncol =1 )
```

# Statistical Comparison of AUCs

```{r}
calculate_z_p_value <- function(mean1, mean2, se1, se2) {
  #' Calculates the Z-statistic and two-tailed p-value for the difference of two means.
  #'
  #' Args:
  #'   mean1 (numeric): The mean of the first group.
  #'   mean2 (numeric): The mean of the second group.
  #'   se1 (numeric): The standard error of the mean of the first group.
  #'   se2 (numeric): The standard error of the mean of the second group.
  #'
  #' Returns:
  #'   A list containing:
  #'     z_statistic (numeric): The calculated Z-statistic.
  #'     p_value (numeric): The two-tailed p-value. Returns NA if standard errors are non-positive.
  #'
  if (se1 <= 0 || se2 <= 0) {
    return(list(z_statistic = NA, p_value = NA))
  } else {
    z_statistic <- (mean1 - mean2) / sqrt(se1^2 + se2^2)
    p_value <- 2 * pnorm(abs(z_statistic), lower.tail = FALSE)
    return(list(z_statistic = z_statistic, p_value = p_value))
  }
}
```

```{r}
auc_comp_sep <- auc_comp_sep %>%
  mutate(se = (upperCI - lowerCI) / (2 *1.96)) %>%
  filter(!is.na(se))

final_results <- data.table()
data <- auc_comp_sep
for (tx in treatment){
  print(tx)
  for (phx in unique(data$outcome)){
    print(phx)
    df <- data %>%
      filter(treatment == tx) %>% filter(outcome == phx) %>% mutate(group = paste0(method, "_", received_tx))
    GROUPS = unique(df$group)
    COMPARISONS_MATRIX <- combn(GROUPS, 2)
    COMPARISONS <- lapply(1:ncol(COMPARISONS_MATRIX), function(i) COMPARISONS_MATRIX[,i])
     # Initialize an empty list to store the results
    for (comparison in COMPARISONS) {
      group_name1 <- comparison[1]
      group_name2 <- comparison[2]
    
      # Get the data for group 1
      data_group1 <- df%>%
        filter(group == group_name1)
      # Get the data for group 2
      data_group2 <- df %>%
        filter(group == group_name2)
      # Extract mean (auc) and standard error (se) for group 1
      mean1 <- data_group1$auc[1] # Assuming only one row per group in df_input
      se1 <- data_group1$se[1]
      # Extract mean (auc) and standard error (se) for group 2
      mean2 <- data_group2$auc[1] # Assuming only one row per group in df_input
      se2 <- data_group2$se[1]
      # Calculate Z-score and p-value using your function
      z_p_result <- calculate_z_p_value(mean1, mean2, se1, se2)
      # Store the results for this comparison
      new_row<- data.frame(
        group1 = group_name1,
        group2 = group_name2,
        difference_in_means = mean1 - mean2,
        z_statistic = z_p_result$z_statistic,
        p_value = z_p_result$p_value,
        treatment = tx,
        outcome = phx
      )
      final_results <- rbind(final_results, new_row)
    }
  }
}
```

```{r}
sig_results_auc <- final_results %>% filter(p_value < 0.05) %>% mutate(group=paste0(treatment, "_", outcome))

auc_comp_sep_sig <- auc_comp_sep %>% mutate(group=paste0(treatment, "_", outcome)) %>% filter( group %in% sig_results_auc$group)
ggplot(auc_comp_sep_sig, aes(x = auc, y = outcome, fill = interaction(method, received_tx))) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +  # Horizontal bars with borders
  geom_errorbar(aes(xmin = lowerCI, xmax = upperCI), 
                width = 0.3,  # Width of the error bar ends
                position = position_dodge(width = 0.9)) +  # Dodge to align with bars
  geom_text(aes(label = round(auc, 3), x = auc - .1),  # Add labels slightly to the right
        position = position_dodge(width = 0.9),    # Align with bars
        size = 4,                                  # Adjust text size
        hjust = 1) +
  labs(
    x = "AUC",
    y = "Phenotype",
    title = "AUC for Phenotypes by Method"
  ) +
  theme_bw(14) +  # Clean theme
  theme(
    axis.text.y = element_text(size = 10),  # Adjust y-axis text size
    axis.title = element_text(size = 12),  # Adjust axis title size
    legend.position = "right"  # Place legend on the right
  ) +
  scale_fill_npg() + facet_wrap(~treatment, ncol =1, scales = "free_y")

```


# Pull Metrics for PGS

```{r}

cox_results <- readRDS("/bd-fs-mnt/TenantShare/exthome/home/l578311/cvd/pgs_revised/data/results_data/all_results_norounding.rds")

```

# Debugging AUC issue.

It looks like I need to debug why I am not getting the correct output for some of the PRSMixPlus results...
```{r}
pheno_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_diabetes_phenotype.txt"
covar_file <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/phenotypes/EUR_inc_diabetes_covariates.txt"

treatment <- NA
categorical_vars <- c("bl_inc_cat2", "bl_educlvl_5cat", "bl_meno_status", "smoke_status_6m", "prev_anycmd","prev_anycvd", "rad_tx_yn", "anthra_yn", "tras_yn", "horm_yn")

categorical_vars<- setdiff(categorical_vars, treatment)
# Continuous variables (e.g., age, BMI, counts)
continuous_vars <- c( "dxage", "bmi_new",  "pc_util_count")

outcomes <- c("inc_anycmd", "inc_hyperten", "inc_diabetes", "inc_dyslipid", "inc_anycvd", "inc_seriouscvd", "inc_arrhythmia", "inc_hf", "inc_ischemic", "inc_stroke", "inc_vte", "inc_cvddeath")
cat_def <- c("M","M", "M","M", "V","V","V","V", "V", "V","V", "V")
dict <-setNames(cat_def, outcomes)

pheno_name <- "inc_diabetes"

if(dict[[pheno_name]]=="M"){
  categorical_vars <- categorical_vars[categorical_vars !="prev_anycmd"]
}else{
  categorical_vars <-  categorical_vars[categorical_vars !="prev_anycvd"]
}

covariates <- c(categorical_vars, continuous_vars)

scores_file_vector <-fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_diabetes_file_paths.txt", header =F)$V1 

pgs_list1 <- fread("/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_diabetes_PGS_list.txt", header = F)
num_pgs <- nrow(pgs_list1)

pheno_file <- pheno_file
covariate_file <- covar_file
score_files_list = scores_file_vector 
trait_specific_score_file = "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_diabetes_PGS_list.txt"
isbinary <- TRUE
out <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/test_out"
allPGS_list = NULL
metascore = NULL 
liabilityR2 = TRUE
IID_pheno = "IID" 
covar_list = covariates
cat_covar_list = categorical_vars
ncores = 1
is_extract_adjSNPeff = F
train_size_list = NULL
training_result_file = NULL
power_thres_list = c(0.95) 
pval_thres_list = (.05/7)
nfold_cv = 3
read_pred_training = FALSE
read_pred_testing = FALSE
debug = F 

```


```{r}
options(datatable.fread.datatable = FALSE)
writeLines("--- Reading covariate data ---")
basic_data = fread(covariate_file)
basic_data = basic_data[, c(IID_pheno, covar_list)]
if (length(cat_covar_list) > 0) {
  writeLines("Generating dummy variables from categorical variables")
  basic_data_list = make_dummy_columns(basic_data, cat_covar_list)
  covar_list = covar_list[-match(cat_covar_list, covar_list)]
  covar_list = c(covar_list, basic_data_list[[2]])
  basic_data = basic_data_list[[1]]
}
writeLines("--- Reading all polygenic risk scores ---")
writeLines("Please make sure the column names for the scores ends with _SUM")
all_scores = NULL
for (score_file_i in 1:length(score_files_list)) {
  print(score_file_i)
  score_file = score_files_list[score_file_i]
  dd = fread(score_file)
  idx = which(endsWith(colnames(dd), "_SUM") & colnames(dd) != 
    "NAMED_ALLELE_DOSAGE_SUM")
  idx2 = which(colnames(dd) == "IID")
  dd_sub = dd[, c(idx2, idx)]
  print(dim(dd_sub))
  if (is.null(all_scores)) {
    all_scores = dd_sub
  }
  else all_scores = merge(all_scores, dd_sub, by = "IID")
}
colnames(all_scores)[2:ncol(all_scores)] = substring(colnames(all_scores)[2:ncol(all_scores)], 
  1, nchar(colnames(all_scores)[2:ncol(all_scores)]) - 
    4)
if (!is.null(allPGS_list)) {
    pgs_extract = intersect(colnames(all_scores[2:ncol(all_scores)]), 
      allPGS_list)
    all_scores = all_scores[, c(1, match(pgs_extract, colnames(all_scores)))]
}
score_names = colnames(all_scores)[2:ncol(all_scores)]
pgs_list = NULL
for (ff_i in 1:length(trait_specific_score_file)) {
  writeLines(paste0("Reading: ", trait_specific_score_file[ff_i]))
  pgs_list_tmp = fread(trait_specific_score_file[ff_i], 
    header = F)[, 1]
  pgs_list = c(pgs_list, pgs_list_tmp)
}
pgs_list = intersect(pgs_list, colnames(all_scores))
pheno = fread(pheno_file)
idx = which(colnames(pheno) %in% c(IID_pheno, pheno_name))
pheno = pheno[, idx]
colnames(pheno) = c("IID", "trait")
```

#### Merging Phenotype and PRS Files

```{r}
writeLines("--- Merging Phenotype and PRS files ---")
pheno_prs = merge(pheno, all_scores, by = "IID")
pheno_prs_cov = merge(pheno_prs, basic_data, by.x = "IID", 
  by.y = IID_pheno)
pheno_prs_cov = pheno_prs_cov[which(!is.na(pheno_prs_cov$trait)), 
  ]
irnt = function(x) return(qnorm((rank(x, na.last = "keep") - 
  0.5)/sum(!is.na(x))))
rr = function(x, d = 3) round(x, d)
null_train_size_list = F
if (is.null(train_size_list)) {
  writeLines("--- Selecting 80% as training data ---")
  train_size_list = floor(0.8 * nrow(pheno_prs_cov))
  null_train_size_list = T
} else {
  writeLines("Using custom data split!")
}
out_save = out
```

```{r}

train_size = train_size_list[1]
writeLines(paste("--- Using ", train_size, " individuals for training sample ---"))
if (!null_train_size_list) 
  out = paste0(out_save, "_train.", train_size)
if (isbinary) 
  fwrite(as.data.frame(table(pheno_prs_cov$trait)), 
    paste0(out, "_case_counts.txt"), row.names = F, 
    sep = "\t", quote = F)
set.seed(1)
train_idx = sample(1:nrow(pheno_prs_cov), train_size)
train_df = pheno_prs_cov[train_idx, ]
test_df = pheno_prs_cov[-train_idx, ]
if (!isbinary) {
  train_df$trait = irnt(train_df$trait)
  test_df$trait = irnt(test_df$trait)
}

cc = apply(train_df[, covar_list], 2, function(x) length(unique(x)))
covar_list1 = covar_list[-which(cc <= 5)]
for (i in covar_list1) train_df[i] = as.numeric(scale(train_df[i]))
for (i in covar_list1) test_df[i] = as.numeric(scale(test_df[i]))
fwrite(train_df[, c(1, 2)], paste0(out, "_train_df.txt"), 
  row.names = F, quote = F, sep = "\t")
fwrite(test_df[, c(1, 2)], paste0(out, "_test_df.txt"), 
  row.names = F, quote = F, sep = "\t")
writeLines("--- Evaluating PRS in training set ---")
if (!read_pred_training | is.null(training_result_file)) {
  training_file = paste0(out, "_train_allPRS.txt")
  read_pred_training_1 = (read_pred_training & file.exists(training_file))
}else {
  training_file = training_result_file
  read_pred_training_1 = read_pred_training & all(file.exists(training_file))
  if (!read_pred_training_1) {
    writeLines("Declared:")
    writeLines(paste0("read_pred_training = ", read_pred_training))
    writeLines(paste0("training_result_file = ", 
      paste(training_result_file, collapse = ";")))
    stop("Reading existing training result but there is at least one missing training result file. Please check parameter all files available in <training_result_file> or set read_pred_testing=F and training_result_file=NULL")
  }
  writeLines(paste0("Reading all files in training_result_file"))
}
if (!read_pred_training_1) {
  sumscore = apply(as.data.frame(train_df[, score_names]), 
    2, var)
  idx = which(sumscore == 0)
  if (length(idx) > 0) 
    train_df = train_df[, -match(names(idx), colnames(train_df))]
  pgs_list_all = colnames(train_df)
  pgs_list_all = pgs_list_all[which(pgs_list_all %in% 
    colnames(all_scores)[2:ncol(all_scores)])]
  pred_acc_train_allPGS_summary = eval_multiple_PRS(train_df, 
    pgs_list_all, covar_list, liabilityR2, alpha = 0.05, 
    isbinary = isbinary)
  head(pred_acc_train_allPGS_summary)
  fwrite(pred_acc_train_allPGS_summary, training_file, 
    row.names = F, sep = "\t", quote = F)
}else {
  pred_acc_train_allPGS_summary = NULL
  for (file_i in 1:length(training_file)) {
    writeLines(paste0("Reading training file: ", 
      training_file[file_i]))
    pred_acc_train_allPGS_summary_tmp = fread(training_file[file_i])
    pred_acc_train_allPGS_summary = rbind(pred_acc_train_allPGS_summary, 
      pred_acc_train_allPGS_summary_tmp)
  }
  pred_acc_train_allPGS_summary = pred_acc_train_allPGS_summary[!duplicated(pred_acc_train_allPGS_summary$pgs), 
    ]
}
```

```{r}
pred_acc_train_allPGS_summary = as.data.frame(pred_acc_train_allPGS_summary)
pred_acc_train_trait_summary = pred_acc_train_allPGS_summary
pred_acc_train_trait_summary = pred_acc_train_trait_summary[order(as.numeric(pred_acc_train_trait_summary$pval), 
  decreasing = F), ]
head(pred_acc_train_trait_summary)
pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary %>% 
  filter(pgs %in% pgs_list)
pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary1[order(as.numeric(pred_acc_train_allPGS_summary1$R2), 
  decreasing = T), ]
bestPRS = pred_acc_train_allPGS_summary1[1, 1]
writeLines(paste0("The best single trait-specific score in the training set is ", 
  bestPRS))
bestPRS_acc = eval_single_PRS(test_df, pheno = "trait", 
  prs_name = bestPRS, covar_list = covar_list, liabilityR2, 
  alpha = 0.05, isbinary = isbinary)
fwrite(bestPRS_acc, paste0(out, "_best_acc.txt"), row.names = F, 
  sep = "\t", quote = F)
writeLines("--- Evaluating PRS in testing set ---")
testing_file = paste0(out, "_test_allPRS.txt")
read_pred_testing_1 = (read_pred_testing & file.exists(testing_file))
if (!read_pred_testing_1) {
  sumscore = apply(test_df[, score_names], 2, var)
  idx = which(sumscore == 0)
  if (length(idx) > 0) 
    test_df = test_df[, -match(names(idx), colnames(test_df))]
  pred_acc_test_trait = eval_multiple_PRS(test_df, 
    pgs_list, covar_list, liabilityR2, alpha = 0.05, 
    isbinary = isbinary)
  pred_acc_test_trait_summary = pred_acc_test_trait
  pred_acc_test_trait_summary = pred_acc_test_trait_summary[order(as.numeric(pred_acc_test_trait_summary$pval), 
    decreasing = F), ]
  print(head(pred_acc_test_trait_summary))
  fwrite(pred_acc_test_trait_summary, testing_file, 
    row.names = F, sep = "\t", quote = F)
} else {
      writeLines(paste0("Reading testing file: ", testing_file))
      pred_acc_test_trait_summary = fread(testing_file)
}
```


```{r}
x_train = (train_df %>% select(all_of(c(bestPRS, 
        covar_list)), -trait))
y_train = as.vector(train_df$trait)
train_data = data.frame(x_train, trait = y_train)
x_test = (test_df %>% select(all_of(c(bestPRS, covar_list)), 
  -trait))
y_test = as.vector(test_df$trait)
test_data = data.frame(x_test, trait = y_test)
train_tmp = train_data[, c("trait", c(bestPRS, covar_list))]
train_tmp$trait = as.factor(train_tmp$trait)
formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, 
  collapse = "+")))
formula_bestPRS = as.formula(paste0("trait ~ scale(", 
  bestPRS, ")+", paste0(covar_list, collapse = "+")))
ctrl = trainControl(method = "repeatedcv", allowParallel = TRUE, 
  number = nfold_cv, returnData = FALSE, trim = T, 
  verboseIter = T)
cl = makePSOCKcluster(ncores)
cl = makeCluster(ncores)
registerDoParallel(cl)
set.seed(123)
model_prsmix_null = train(formula_null, data = train_tmp, 
  method = "glmnet", trControl = ctrl, family = "binomial", 
  tuneLength = 50, verbose = T)
set.seed(123)
model_prsmix_bestPRS = train(formula_bestPRS, data = train_tmp, 
  method = "glmnet", trControl = ctrl, family = "binomial", 
  tuneLength = 50, verbose = T)
stopCluster(cl)
```
      
```{r}
test_data1 = test_data
test_pred = predict(model_prsmix_null, test_data1, 
  type = "prob")[, 2]
auc_ci = ci.auc(test_data1$trait, test_pred)
auc_out = data.frame(method = "null_model", auc = auc_ci[2], 
  lowerCI = auc_ci[1], upperCI = auc_ci[3])
fwrite(auc_out, paste0(out, "_auc_NULL.txt"), row.names = F, 
  sep = "\t", quote = F)
test_pred = predict(model_prsmix_bestPRS, test_data1, 
  type = "prob")[, 2]
auc_ci = ci.auc(test_data1$trait, test_pred)
auc_out = data.frame(method = "bestPGS", auc = auc_ci[2], 
  lowerCI = auc_ci[1], upperCI = auc_ci[3])
fwrite(auc_out, paste0(out, "_auc_BestPGS.txt"), 
  row.names = F, sep = "\t", quote = F)
model1 = glm(formula_bestPRS, data = test_data1, 
  family = "binomial")
model1s = summary(model1)
mm = exp(model1s$coefficients[2, 1])
ll = exp(model1s$coefficients[2, 1] - 1.97 * model1s$coefficients[2, 
  2])
uu = exp(model1s$coefficients[2, 1] + 1.97 * model1s$coefficients[2, 
  2])
pval = format.pval(model1s$coefficients[2, 4])
writeLines(paste0("OR = ", rr(mm), " (", rr(ll), 
  "-", rr(uu), "); P-value=", pval))
fwrite(data.frame(mm, ll, uu, pval), paste0(out, 
  "_OR_BestPGS.txt"), row.names = F, sep = "\t", 
  quote = F)
```


```{r}
pred_acc_train_trait_summary = pred_acc_train_allPGS_summary %>% 
      filter(pgs %in% pgs_list)
pred_acc_test_trait_summary_out = pred_acc_test_trait_summary
writeLines("--- Iterating power and p-value parameters --- ")
```


```{r}
power_thres <- 0.95
pval_thres <- .05

writeLines(paste0("P = ", pval_thres))
writeLines(paste0("Power = ", power_thres))
writeLines("PRSmix:")
topprs = pred_acc_train_trait_summary %>% filter(pval <= 
  pval_thres & power >= power_thres)
head(topprs)
topprs = topprs$pgs
topprs = intersect(topprs, colnames(train_df))
print(length(topprs))
start_time = Sys.time()
```

```{r}
if (!isbinary) {
  x_train = (train_df %>% select(all_of(c(topprs, 
    covar_list)), -trait))
  if (length(topprs) > 1) 
    sd_train = apply(as.data.frame(x_train[, 
      topprs]), 2, sd, na.rm = T)
  x_train[, topprs] = scale(x_train[, topprs])
  y_train = as.vector(train_df$trait)
  train_data = data.frame(x_train, trait = y_train)
  x_test = (test_df %>% select(all_of(c(topprs, 
    covar_list)), -trait))
  y_test = as.vector(test_df$trait)
  test_data = data.frame(x_test, trait = y_test)
  formula = as.formula(paste0("trait ~ ", paste0(topprs, 
    collapse = "+"), "+", paste0(covar_list, 
    collapse = "+")))
  train_tmp = train_data[, c("trait", topprs, 
    covar_list)]
  writeLines("Header of training data")
  print(head(train_tmp))
  if (length(topprs) == 1) {
    ww = ww_raw = c(1)
    names(ww) = names(ww_raw) = topprs
  }
  else {
    ctrl = trainControl(method = "repeatedcv", 
      allowParallel = TRUE, number = nfold_cv, 
      verboseIter = T)
    cl = makePSOCKcluster(ncores)
    registerDoParallel(cl)
    set.seed(123)
    model_prsmix = train(formula, data = train_tmp, 
      method = "glmnet", trControl = ctrl, tuneLength = 50, 
      verbose = T)
    stopCluster(cl)
    model_prsmix$bestTune
    ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[, 
      1][-1]
    ww = ww[which(!names(ww) %in% covar_list)]
    ww_raw = ww
    if (all(ww == 0)) {
      writeLines("No weight for PRS")
      ww = c(1)
      names(ww) = bestPRS_acc$pgs
      topprs = bestPRS_acc$pgs
    }
    else {
      ww = ww/sd_train[match(names(ww), names(sd_train))]
    }
  }
  test_df1 = cbind(test_data, IID = test_df$IID)
  test_df1$newprs = as.matrix(test_df1[, topprs]) %*% 
    as.vector(ww)
  res_lm1 = eval_single_PRS(test_df1, pheno = "trait", 
    prs_name = "newprs", covar_list = covar_list, 
    liabilityR2 = liabilityR2, alpha = pval_thres, 
    isbinary = isbinary)
  res_lm1$pgs = "PRSmix"
  res_lm1
}else {
  x_train = (train_df %>% select(all_of(c(topprs, 
    covar_list)), -trait))
  if (length(topprs) > 1) 
    sd_train = apply(as.data.frame(x_train[, 
      topprs]), 2, sd, na.rm = T)
  x_train[, topprs] = scale(x_train[, topprs])
  y_train = as.vector(train_df$trait)
  train_data = data.frame(x_train, trait = y_train)
  x_test = (test_df %>% select(all_of(c(topprs, 
    covar_list)), -trait))
  y_test = as.vector(test_df$trait)
  test_data = data.frame(x_test, trait = y_test)
  formula = as.formula(paste0("trait ~ ", paste0(topprs, 
    collapse = "+"), "+", paste0(covar_list, 
    collapse = "+")))
  train_tmp = train_data[, c("trait", topprs, 
    covar_list)]
  train_tmp$trait = as.factor(train_tmp$trait)
  writeLines("Header of training data")
  print(head(train_tmp))
  if (length(topprs) == 1) {
    ww = ww_raw = c(1)
    names(ww) = names(ww_raw) = topprs
  } else {
    ctrl = trainControl(method = "repeatedcv", 
      allowParallel = TRUE, number = nfold_cv, 
      verboseIter = T)
    cl = makePSOCKcluster(ncores)
    registerDoParallel(cl)
    set.seed(123)
    model_prsmix = train(formula, data = train_tmp, 
      method = "glmnet", trControl = ctrl, family = "binomial", 
      tuneLength = 50, verbose = T)
    stopCluster(cl)
    model_prsmix$bestTune
    ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[, 
      1][-1]
    ww_raw = ww
    if (all(ww == 0)) {
      writeLines("No weight for PRS")
      ww = c(1)
      names(ww) = bestPRS_acc$pgs
      topprs = bestPRS_acc$pgs
    }    else {
      ww = ww[which(!names(ww) %in% covar_list)]
      ww = ww/sd_train[match(names(ww), names(sd_train))]
    }
    if (debug) {
      print(ww)
    }
    test_data1 = test_data
    test_data1[, topprs] = as.numeric(scale(test_data[, 
      topprs]))
    test_pred = predict(model_prsmix, test_data1, 
      type = "prob")[, 2]
    auc_ci = ci.auc(test_data1$trait, test_pred)
    auc_out = data.frame(method = "prsmix", 
      auc = auc_ci[2], lowerCI = auc_ci[1], 
      upperCI = auc_ci[3])
    fwrite(auc_out, paste0(out, "_power.", power_thres, 
      "_pthres.", pval_thres, "_auc_PRSmix.txt"), 
      row.names = F, sep = "\t", quote = F)
  }
}
```

```{r full_script_debug}
options(datatable.fread.datatable = FALSE)
writeLines("--- Reading covariate data ---")
basic_data = fread(covariate_file)
basic_data = basic_data[, c(IID_pheno, covar_list)]
if (length(cat_covar_list) > 0) {
  writeLines("Generating dummy variables from categorical variables")
  basic_data_list = make_dummy_columns(basic_data, cat_covar_list)
  covar_list = covar_list[-match(cat_covar_list, covar_list)]
  covar_list = c(covar_list, basic_data_list[[2]])
  basic_data = basic_data_list[[1]]
}
writeLines("--- Reading all polygenic risk scores ---")
writeLines("Please make sure the column names for the scores ends with _SUM")
all_scores = NULL
for (score_file_i in 1:length(score_files_list)) {
  print(score_file_i)
  score_file = score_files_list[score_file_i]
  dd = fread(score_file)
  idx = which(endsWith(colnames(dd), "_SUM") & colnames(dd) != 
    "NAMED_ALLELE_DOSAGE_SUM")
  idx2 = which(colnames(dd) == "IID")
  dd_sub = dd[, c(idx2, idx)]
  print(dim(dd_sub))
  if (is.null(all_scores)) {
    all_scores = dd_sub
  }
  else all_scores = merge(all_scores, dd_sub, by = "IID")
}
colnames(all_scores)[2:ncol(all_scores)] = substring(colnames(all_scores)[2:ncol(all_scores)], 
  1, nchar(colnames(all_scores)[2:ncol(all_scores)]) - 
    4)
if (!is.null(allPGS_list)) {
  pgs_extract = intersect(colnames(all_scores[2:ncol(all_scores)]), 
    allPGS_list)
  all_scores = all_scores[, c(1, match(pgs_extract, colnames(all_scores)))]
}
score_names = colnames(all_scores)[2:ncol(all_scores)]
pgs_list = NULL
for (ff_i in 1:length(trait_specific_score_file)) {
  writeLines(paste0("Reading: ", trait_specific_score_file[ff_i]))
  pgs_list_tmp = fread(trait_specific_score_file[ff_i], 
    header = F)[, 1]
  pgs_list = c(pgs_list, pgs_list_tmp)
}
pgs_list = intersect(pgs_list, colnames(all_scores))
pheno = fread(pheno_file)
idx = which(colnames(pheno) %in% c(IID_pheno, pheno_name))
pheno = pheno[, idx]
colnames(pheno) = c("IID", "trait")
writeLines("--- Merging Phenotype and PRS files ---")
pheno_prs = merge(pheno, all_scores, by = "IID")
pheno_prs_cov = merge(pheno_prs, basic_data, by.x = "IID", 
  by.y = IID_pheno)
pheno_prs_cov = pheno_prs_cov[which(!is.na(pheno_prs_cov$trait)), 
  ]
irnt = function(x) return(qnorm((rank(x, na.last = "keep") - 
  0.5)/sum(!is.na(x))))
rr = function(x, d = 3) round(x, d)
null_train_size_list = F
if (is.null(train_size_list)) {
  writeLines("--- Selecting 80% as training data ---")
  train_size_list = floor(0.8 * nrow(pheno_prs_cov))
  null_train_size_list = T
}else {
  writeLines("Using custom data split!")
}
out_save = out
for (train_size in train_size_list) {
  train_size = train_size_list[1]
  writeLines(paste("--- Using ", train_size, " individuals for training sample ---"))
  if (!null_train_size_list) 
    out = paste0(out_save, "_train.", train_size)
  if (isbinary) 
    fwrite(as.data.frame(table(pheno_prs_cov$trait)), 
      paste0(out, "_case_counts.txt"), row.names = F, 
      sep = "\t", quote = F)
  set.seed(1)
  train_idx = sample(1:nrow(pheno_prs_cov), train_size)
  train_df = pheno_prs_cov[train_idx, ]
  test_df = pheno_prs_cov[-train_idx, ]
  if (!isbinary) {
    train_df$trait = irnt(train_df$trait)
    test_df$trait = irnt(test_df$trait)
  }
  cc = apply(train_df[, covar_list], 2, function(x) length(unique(x)))
  covar_list1 = covar_list[-which(cc <= 5)]
  for (i in covar_list1) train_df[i] = as.numeric(scale(train_df[i]))
  for (i in covar_list1) test_df[i] = as.numeric(scale(test_df[i]))
  fwrite(train_df[, c(1, 2)], paste0(out, "_train_df.txt"), 
    row.names = F, quote = F, sep = "\t")
  fwrite(test_df[, c(1, 2)], paste0(out, "_test_df.txt"), 
    row.names = F, quote = F, sep = "\t")
  writeLines("--- Evaluating PRS in training set ---")
  if (!read_pred_training | is.null(training_result_file)) {
    training_file = paste0(out, "_train_allPRS.txt")
    read_pred_training_1 = (read_pred_training & file.exists(training_file))
  }
  else {
    training_file = training_result_file
    read_pred_training_1 = read_pred_training & all(file.exists(training_file))
    if (!read_pred_training_1) {
      writeLines("Declared:")
      writeLines(paste0("read_pred_training = ", read_pred_training))
      writeLines(paste0("training_result_file = ", 
        paste(training_result_file, collapse = ";")))
      stop("Reading existing training result but there is at least one missing training result file. Please check parameter all files available in <training_result_file> or set read_pred_testing=F and training_result_file=NULL")
    }
    writeLines(paste0("Reading all files in training_result_file"))
  }
  if (!read_pred_training_1) {
    sumscore = apply(as.data.frame(train_df[, score_names]), 
      2, var)
    idx = which(sumscore == 0)
    if (length(idx) > 0) 
      train_df = train_df[, -match(names(idx), colnames(train_df))]
    pgs_list_all = colnames(train_df)
    pgs_list_all = pgs_list_all[which(pgs_list_all %in% 
      colnames(all_scores)[2:ncol(all_scores)])]
    pred_acc_train_allPGS_summary = eval_multiple_PRS(train_df, 
      pgs_list_all, covar_list, liabilityR2, alpha = 0.05, 
      isbinary = isbinary)
    head(pred_acc_train_allPGS_summary)
    fwrite(pred_acc_train_allPGS_summary, training_file, 
      row.names = F, sep = "\t", quote = F)
  }
  else {
    pred_acc_train_allPGS_summary = NULL
    for (file_i in 1:length(training_file)) {
      writeLines(paste0("Reading training file: ", 
        training_file[file_i]))
      pred_acc_train_allPGS_summary_tmp = fread(training_file[file_i])
      pred_acc_train_allPGS_summary = rbind(pred_acc_train_allPGS_summary, 
        pred_acc_train_allPGS_summary_tmp)
    }
    pred_acc_train_allPGS_summary = pred_acc_train_allPGS_summary[!duplicated(pred_acc_train_allPGS_summary$pgs), 
      ]
  }
  pred_acc_train_allPGS_summary = as.data.frame(pred_acc_train_allPGS_summary)
  pred_acc_train_trait_summary = pred_acc_train_allPGS_summary
  pred_acc_train_trait_summary = pred_acc_train_trait_summary[order(as.numeric(pred_acc_train_trait_summary$pval), 
    decreasing = F), ]
  head(pred_acc_train_trait_summary)
  pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary %>% 
    filter(pgs %in% pgs_list)
  pred_acc_train_allPGS_summary1 = pred_acc_train_allPGS_summary1[order(as.numeric(pred_acc_train_allPGS_summary1$R2), 
    decreasing = T), ]
  bestPRS = pred_acc_train_allPGS_summary1[1, 1]
  writeLines(paste0("The best single trait-specific score in the training set is ", 
    bestPRS))
  bestPRS_acc = eval_single_PRS(test_df, pheno = "trait", 
    prs_name = bestPRS, covar_list = covar_list, liabilityR2, 
    alpha = 0.05, isbinary = isbinary)
  fwrite(bestPRS_acc, paste0(out, "_best_acc.txt"), row.names = F, 
    sep = "\t", quote = F)
  writeLines("--- Evaluating PRS in testing set ---")
  testing_file = paste0(out, "_test_allPRS.txt")
  read_pred_testing_1 = (read_pred_testing & file.exists(testing_file))
  if (!read_pred_testing_1) {
    sumscore = apply(test_df[, score_names], 2, var)
    idx = which(sumscore == 0)
    if (length(idx) > 0) 
      test_df = test_df[, -match(names(idx), colnames(test_df))]
    pred_acc_test_trait = eval_multiple_PRS(test_df, 
      pgs_list, covar_list, liabilityR2, alpha = 0.05, 
      isbinary = isbinary)
    pred_acc_test_trait_summary = pred_acc_test_trait
    pred_acc_test_trait_summary = pred_acc_test_trait_summary[order(as.numeric(pred_acc_test_trait_summary$pval), 
      decreasing = F), ]
    print(head(pred_acc_test_trait_summary))
    fwrite(pred_acc_test_trait_summary, testing_file, 
      row.names = F, sep = "\t", quote = F)
  }
  else {
    writeLines(paste0("Reading testing file: ", testing_file))
    pred_acc_test_trait_summary = fread(testing_file)
  }
  if (isbinary) {
    x_train = (train_df %>% select(all_of(c(bestPRS, 
      covar_list)), -trait))
    y_train = as.vector(train_df$trait)
    train_data = data.frame(x_train, trait = y_train)
    x_test = (test_df %>% select(all_of(c(bestPRS, covar_list)), 
      -trait))
    y_test = as.vector(test_df$trait)
    test_data = data.frame(x_test, trait = y_test)
    train_tmp = train_data[, c("trait", c(bestPRS, covar_list))]
    train_tmp$trait = as.factor(train_tmp$trait)
    formula_null = as.formula(paste0("trait ~ ", paste0(covar_list, 
      collapse = "+")))
    formula_bestPRS = as.formula(paste0("trait ~ scale(", 
      bestPRS, ")+", paste0(covar_list, collapse = "+")))
    ctrl = trainControl(method = "repeatedcv", allowParallel = TRUE, 
      number = nfold_cv, returnData = FALSE, trim = T, 
      verboseIter = T)
    cl = makePSOCKcluster(ncores)
    cl = makeCluster(ncores)
    registerDoParallel(cl)
    set.seed(123)
    model_prsmix_null = train(formula_null, data = train_tmp, 
      method = "glmnet", trControl = ctrl, family = "binomial", 
      tuneLength = 50, verbose = T)
    set.seed(123)
    model_prsmix_bestPRS = train(formula_bestPRS, data = train_tmp, 
      method = "glmnet", trControl = ctrl, family = "binomial", 
      tuneLength = 50, verbose = T)
    stopCluster(cl)
    test_data1 = test_data
    test_pred = predict(model_prsmix_null, test_data1, 
      type = "prob")[, 2]
    auc_ci = ci.auc(test_data1$trait, test_pred)
    auc_out = data.frame(method = "null_model", auc = auc_ci[2], 
      lowerCI = auc_ci[1], upperCI = auc_ci[3])
    fwrite(auc_out, paste0(out, "_auc_NULL.txt"), row.names = F, 
      sep = "\t", quote = F)
    test_pred = predict(model_prsmix_bestPRS, test_data1, 
      type = "prob")[, 2]
    auc_ci = ci.auc(test_data1$trait, test_pred)
    auc_out = data.frame(method = "bestPGS", auc = auc_ci[2], 
      lowerCI = auc_ci[1], upperCI = auc_ci[3])
    fwrite(auc_out, paste0(out, "_auc_BestPGS.txt"), 
      row.names = F, sep = "\t", quote = F)
    model1 = glm(formula_bestPRS, data = test_data1, 
      family = "binomial")
    model1s = summary(model1)
    mm = exp(model1s$coefficients[2, 1])
    ll = exp(model1s$coefficients[2, 1] - 1.97 * model1s$coefficients[2, 
      2])
    uu = exp(model1s$coefficients[2, 1] + 1.97 * model1s$coefficients[2, 
      2])
    pval = format.pval(model1s$coefficients[2, 4])
    writeLines(paste0("OR = ", rr(mm), " (", rr(ll), 
      "-", rr(uu), "); P-value=", pval))
    fwrite(data.frame(mm, ll, uu, pval), paste0(out, 
      "_OR_BestPGS.txt"), row.names = F, sep = "\t", 
      quote = F)
  }
  pred_acc_train_trait_summary = pred_acc_train_allPGS_summary %>% 
    filter(pgs %in% pgs_list)
  pred_acc_test_trait_summary_out = pred_acc_test_trait_summary
  writeLines("--- Iterating power and p-value parameters --- ")
  for (power_thres in power_thres_list) for (pval_thres in pval_thres_list) {
    writeLines(paste0("P = ", pval_thres))
    writeLines(paste0("Power = ", power_thres))
    writeLines("PRSmix:")
    topprs = pred_acc_train_trait_summary %>% filter(pval <= 
      pval_thres & power >= power_thres)
    head(topprs)
    topprs = topprs$pgs
    topprs = intersect(topprs, colnames(train_df))
    print(length(topprs))
    start_time = Sys.time()
    if (length(topprs) == 0) {
      print("No high power trait-specific PRS for PRSmix")
      ww_raw = 1
    }
    else {
      if (!isbinary) {
        x_train = (train_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        if (length(topprs) > 1) 
          sd_train = apply(as.data.frame(x_train[, 
            topprs]), 2, sd, na.rm = T)
        x_train[, topprs] = scale(x_train[, topprs])
        y_train = as.vector(train_df$trait)
        train_data = data.frame(x_train, trait = y_train)
        x_test = (test_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        y_test = as.vector(test_df$trait)
        test_data = data.frame(x_test, trait = y_test)
        formula = as.formula(paste0("trait ~ ", paste0(topprs, 
          collapse = "+"), "+", paste0(covar_list, 
          collapse = "+")))
        train_tmp = train_data[, c("trait", topprs, 
          covar_list)]
        writeLines("Header of training data")
        print(head(train_tmp))
        if (length(topprs) == 1) {
          ww = ww_raw = c(1)
          names(ww) = names(ww_raw) = topprs
        }
        else {
          ctrl = trainControl(method = "repeatedcv", 
            allowParallel = TRUE, number = nfold_cv, 
            verboseIter = T)
          cl = makePSOCKcluster(ncores)
          registerDoParallel(cl)
          set.seed(123)
          model_prsmix = train(formula, data = train_tmp, 
            method = "glmnet", trControl = ctrl, tuneLength = 50, 
            verbose = T)
          stopCluster(cl)
          model_prsmix$bestTune
          ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[, 
            1][-1]
          ww = ww[which(!names(ww) %in% covar_list)]
          ww_raw = ww
          if (all(ww == 0)) {
            writeLines("No weight for PRS")
            ww = c(1)
            names(ww) = bestPRS_acc$pgs
            topprs = bestPRS_acc$pgs
          }
          else {
            ww = ww/sd_train[match(names(ww), names(sd_train))]
          }
        }
        test_df1 = cbind(test_data, IID = test_df$IID)
        test_df1$newprs = as.matrix(test_df1[, topprs]) %*% 
          as.vector(ww)
        res_lm1 = eval_single_PRS(test_df1, pheno = "trait", 
          prs_name = "newprs", covar_list = covar_list, 
          liabilityR2 = liabilityR2, alpha = pval_thres, 
          isbinary = isbinary)
        res_lm1$pgs = "PRSmix"
        res_lm1
      }
      else {
        x_train = (train_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        if (length(topprs) > 1) 
          sd_train = apply(as.data.frame(x_train[, 
            topprs]), 2, sd, na.rm = T)
        x_train[, topprs] = scale(x_train[, topprs])
        y_train = as.vector(train_df$trait)
        train_data = data.frame(x_train, trait = y_train)
        x_test = (test_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        y_test = as.vector(test_df$trait)
        test_data = data.frame(x_test, trait = y_test)
        formula = as.formula(paste0("trait ~ ", paste0(topprs, 
          collapse = "+"), "+", paste0(covar_list, 
          collapse = "+")))
        train_tmp = train_data[, c("trait", topprs, 
          covar_list)]
        train_tmp$trait = as.factor(train_tmp$trait)
        writeLines("Header of training data")
        print(head(train_tmp))
        if (length(topprs) == 1) {
          ww = ww_raw = c(1)
          names(ww) = names(ww_raw) = topprs
        }
        else {
          ctrl = trainControl(method = "repeatedcv", 
            allowParallel = TRUE, number = nfold_cv, 
            verboseIter = T)
          cl = makePSOCKcluster(ncores)
          registerDoParallel(cl)
          set.seed(123)
          model_prsmix = train(formula, data = train_tmp, 
            method = "glmnet", trControl = ctrl, family = "binomial", 
            tuneLength = 50, verbose = T)
          stopCluster(cl)
          model_prsmix$bestTune
          ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[, 
            1][-1]
          ww_raw = ww
          if (all(ww == 0)) {
            writeLines("No weight for PRS")
            ww = c(1)
            names(ww) = bestPRS_acc$pgs
            topprs = bestPRS_acc$pgs
          }
          else {
            ww = ww[which(!names(ww) %in% covar_list)]
            ww = ww/sd_train[match(names(ww), names(sd_train))]
          }
          if (debug) {
            print(ww)
          }
          test_data1 = test_data
          test_data1[, topprs] = as.numeric(scale(test_data[, 
            topprs]))
          test_pred = predict(model_prsmix, test_data1, 
            type = "prob")[, 2]
          auc_ci = ci.auc(test_data1$trait, test_pred)
          auc_out = data.frame(method = "prsmix", 
            auc = auc_ci[2], lowerCI = auc_ci[1], 
            upperCI = auc_ci[3])
          fwrite(auc_out, paste0(out, "_power.", power_thres, 
            "_pthres.", pval_thres, "_auc_PRSmix.txt"), 
            row.names = F, sep = "\t", quote = F)
        }
        test_df1 = cbind(test_data, IID = test_df$IID)
        test_df1$newprs = as.matrix(test_df1[, topprs]) %*% 
          as.vector(ww)
        res_lm1 = eval_single_PRS(test_df1, pheno = "trait", 
          prs_name = "newprs", covar_list = covar_list, 
          liabilityR2 = liabilityR2, alpha = pval_thres, 
          isbinary = isbinary)
        res_lm1$pgs = "PRSmix"
        res_lm1
        ff = paste0("trait ~ scale(newprs) + ", paste0(covar_list, 
          collapse = "+"))
        model1 = glm(ff, data = test_df1, family = "binomial")
        model1s = summary(model1)
        mm = exp(model1s$coefficients[2, 1])
        ll = exp(model1s$coefficients[2, 1] - 1.97 * 
          model1s$coefficients[2, 2])
        uu = exp(model1s$coefficients[2, 1] + 1.97 * 
          model1s$coefficients[2, 2])
        pval = format.pval(model1s$coefficients[2, 
          4])
        print(paste0("OR = ", rr(mm), " (", rr(ll), 
          "-", rr(uu), "); P-value=", pval))
        fwrite(data.frame(mm, ll, uu, pval), paste0(out, 
          "_power.", power_thres, "_pthres.", pval_thres, 
          "_OR_PRSmix.txt"), row.names = F, sep = "\t", 
          quote = F)
      }
      fwrite(data.frame(c(topprs), ww), paste0(out, 
        "_power.", power_thres, "_pthres.", pval_thres, 
        "_weight_PRSmix.txt"), sep = "\t", quote = F)
      if (is_extract_adjSNPeff) {
        mixing_weight_file = paste0(out, "_power.", 
          power_thres, "_pthres.", pval_thres, "_weight_PRSmix.txt")
        outfile = paste0(out, "_power.", power_thres, 
          "_pthres.", pval_thres, "_adjSNPeff_PRSmix.txt")
        extract_adjSNPeff(mixing_weight_file, original_beta_files_list, 
          outfile)
      }
      print(res_lm1)
      res_lm1_summary = res_lm1
      res_lm1_summary$pgs = "PRSmix"
      pred_acc_test_trait_summary_out = bind_rows(res_lm1, 
        pred_acc_test_trait_summary)
      head(pred_acc_test_trait_summary_out)
      fwrite(pred_acc_test_trait_summary_out, paste0(out, 
        "_power.", power_thres, "_pthres.", pval_thres, 
        "_test_summary_traitPRS_withPRSmix.txt"), 
        row.names = F, sep = "\t", quote = F)
      prs_out = test_df1 %>% select(IID, newprs)
      colnames(prs_out) = c("IID", "prsmix")
      fwrite(prs_out, paste0(out, "_power.", power_thres, 
        "_pthres.", pval_thres, "_prsmix.txt"), row.names = F, 
        sep = "\t", quote = F)
    }
    end_time = Sys.time()
    timerunning = difftime(end_time, start_time, units = "secs")[[1]]
    timedf = data.frame(pgs = "PRSmix", npgs = length(ww_raw), 
      time = timerunning)
    print(timedf)
    fwrite(timedf, paste0(out, "_power.", power_thres, 
      "_pthres.", pval_thres, "_time_PRSmix.txt"), 
      row.names = F, sep = "\t", quote = F)
    writeLines("PRSmix+:")
    topprs = pred_acc_train_allPGS_summary %>% filter(pval <= 
      pval_thres & power >= power_thres)
    topprs = topprs$pgs
    topprs = intersect(topprs, colnames(train_df))
    print(length(topprs))
    start_time = Sys.time()
    if (length(topprs) == 0) {
      print("No high power trait-specific PRS for PRSmix+")
      ww_raw = 1
    }
    else {
      if (!isbinary) {
        x_train = (train_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        if (length(topprs) > 1) 
          sd_train = apply(as.data.frame(x_train[, 
            topprs]), 2, sd, na.rm = T)
        x_train[, topprs] = scale(x_train[, topprs])
        y_train = as.vector(train_df$trait)
        train_data = data.frame(x_train, trait = y_train)
        x_test = (test_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        y_test = as.vector(test_df$trait)
        test_data = data.frame(x_test, trait = y_test)
        formula = as.formula(paste0("trait ~ ", paste0(topprs, 
          collapse = "+"), "+", paste0(covar_list, 
          collapse = "+")))
        train_tmp = train_data[, c("trait", topprs, 
          covar_list)]
        writeLines("Header of training data")
        print(head(train_tmp))
        if (length(topprs) == 1) {
          ww = c(1)
          names(ww) = topprs
        }
        else {
          ctrl = trainControl(method = "repeatedcv", 
            allowParallel = TRUE, number = nfold_cv, 
            verboseIter = T)
          cl = makePSOCKcluster(ncores)
          registerDoParallel(cl)
          set.seed(123)
          model_prsmix = train(formula, data = train_tmp, 
            method = "glmnet", trControl = ctrl, tuneLength = 50, 
            verbose = T)
          stopCluster(cl)
          model_prsmix$bestTune
          ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[, 
            1][-1]
          ww = ww[which(!names(ww) %in% covar_list)]
          ww_raw = ww
          if (all(ww == 0)) {
            writeLines("No weight for PRS")
            ww = c(1)
            names(ww) = bestPRS_acc$pgs
            topprs = bestPRS_acc$pgs
          }
          else {
            ww = ww/sd_train[match(names(ww), names(sd_train))]
          }
        }
        test_df1 = cbind(test_data, IID = test_df$IID)
        test_df1$newprs = as.matrix(test_df1[, topprs]) %*% 
          as.vector(ww)
        res_lm = eval_single_PRS(test_df1, pheno = "trait", 
          prs_name = "newprs", covar_list = covar_list, 
          liabilityR2 = liabilityR2, alpha = pval_thres, 
          isbinary = isbinary)
        res_lm$pgs = "PRSmix+"
        res_lm
        nonzero_w = names(ww[which(ww != 0)])
      }
      else {
        x_train = (train_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        if (length(topprs) > 1) 
          sd_train = apply(as.data.frame(x_train[, 
            topprs]), 2, sd, na.rm = T)
        x_train[, topprs] = scale(x_train[, topprs])
        y_train = as.vector(train_df$trait)
        train_data = data.frame(x_train, trait = y_train)
        x_test = (test_df %>% select(all_of(c(topprs, 
          covar_list)), -trait))
        y_test = as.vector(test_df$trait)
        test_data = data.frame(x_test, trait = y_test)
        formula = as.formula(paste0("trait ~ ", paste0(topprs, 
          collapse = "+"), "+", paste0(covar_list, 
          collapse = "+")))
        train_tmp = train_data[, c("trait", topprs, 
          covar_list)]
        train_tmp$trait = as.factor(train_tmp$trait)
        writeLines("Header of training data")
        print(head(train_tmp))
        if (length(topprs) == 1) {
          ww = c(1)
          names(ww) = topprs
        }
        else {
          ctrl = trainControl(method = "repeatedcv", 
            allowParallel = TRUE, number = nfold_cv, 
            verboseIter = T)
          cl = makePSOCKcluster(ncores)
          registerDoParallel(cl)
          set.seed(123)
          model_prsmix = train(formula, data = train_tmp, 
            method = "glmnet", trControl = ctrl, family = "binomial", 
            tuneLength = 50, verbose = T)
          stopCluster(cl)
          model_prsmix$bestTune
          ww = coef(model_prsmix$finalModel, model_prsmix$bestTune$lambda)[, 
            1][-1]
          ww = ww[which(!names(ww) %in% covar_list)]
          ww_raw = ww
          if (all(ww == 0)) {
            writeLines("No weight for PRS")
            ww = c(1)
            names(ww) = bestPRS_acc$pgs
            topprs = bestPRS_acc$pgs
          }
          else {
            ww = ww/sd_train[match(names(ww), names(sd_train))]
          }
          test_data1 = test_data
          test_data1[, topprs] = scale(test_data[, 
            topprs])
          test_pred = predict(model_prsmix, test_data1, 
            type = "prob")[, 2]
          auc_ci = ci.auc(test_data1$trait, test_pred)
          auc_out = data.frame(method = "prsmixP", 
            auc = auc_ci[2], lowerCI = auc_ci[1], 
            upperCI = auc_ci[3])
          fwrite(auc_out, paste0(out, "_power.", power_thres, 
            "_pthres.", pval_thres, "_auc_PRSmixPlus.txt"), 
            row.names = F, sep = "\t", quote = F)
        }
        test_df1 = cbind(test_data, IID = test_df$IID)
        test_df1$newprs = as.matrix(test_df1[, topprs]) %*% 
          as.vector(ww)
        res_lm = eval_single_PRS(test_df1, pheno = "trait", 
          "newprs", covar_list, liabilityR2 = liabilityR2, 
          alpha = pval_thres, isbinary = isbinary)
        res_lm$pgs = "PRSmix+"
        res_lm
        nonzero_w = names(ww[which(ww != 0)])
        ff = paste0("trait ~ scale(newprs) + ", paste0(covar_list, 
          collapse = "+"))
        model = glm(ff, data = test_df1, family = "binomial")
        model1s = summary(model)
        mm = exp(model1s$coefficients[2, 1])
        ll = exp(model1s$coefficients[2, 1] - 1.97 * 
          model1s$coefficients[2, 2])
        uu = exp(model1s$coefficients[2, 1] + 1.97 * 
          model1s$coefficients[2, 2])
        pval = format.pval(model1s$coefficients[2, 
          4])
        print(paste0("OR = ", rr(mm), " (", rr(ll), 
          "-", rr(uu), "); P-value=", pval))
        fwrite(data.frame(mm, ll, uu, pval), paste0(out, 
          "_power.", power_thres, "_pthres.", pval_thres, 
          "_OR_PRSmixPlus.txt"), row.names = F, sep = "\t", 
          quote = F)
      }
      print(res_lm)
      pred_acc_test_trait_summary_out = bind_rows(res_lm, 
        pred_acc_test_trait_summary_out)
      head(pred_acc_test_trait_summary_out)
      fwrite(pred_acc_test_trait_summary_out, paste0(out, 
        "_power.", power_thres, "_pthres.", pval_thres, 
        "_test_summary_traitPRS_withPRSmixPlus.txt"), 
        row.names = F, sep = "\t", quote = F)
      prsmixplus = test_df1 %>% select(IID, newprs)
      fwrite(prsmixplus, paste0(out, "_power.", power_thres, 
        "_pthres.", pval_thres, "_prsmixPlus.txt"), 
        row.names = F, sep = "\t", quote = F)
      fwrite(data.frame(topprs, ww), paste0(out, "_power.", 
        power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt"), 
        row.names = F, sep = "\t", quote = F)
      if (is_extract_adjSNPeff) {
        mixing_weight_file = paste0(out, "_power.", 
          power_thres, "_pthres.", pval_thres, "_weight_PRSmixPlus.txt")
        outfile = paste0(out, "_power.", power_thres, 
          "_pthres.", pval_thres, "_adjSNPeff_PRSmixPlus.txt")
        extract_adjSNPeff(mixing_weight_file, original_beta_files_list, 
          outfile)
      }
      fwrite(data.frame(topprs, ww_raw), paste0(out, 
        "_power.", power_thres, "_pthres.", pval_thres, 
        "_weight_raw_PRSmixPlus.txt"), row.names = F, 
        sep = "\t", quote = F)
    }
    end_time = Sys.time()
    timerunning = difftime(end_time, start_time, units = "secs")[[1]]
    timedf = data.frame(pgs = "PRSmix+", npgs = length(ww_raw), 
      time = timerunning)
    print(timedf)
    fwrite(timedf, paste0(out, "_power.", power_thres, 
      "_pthres.", pval_thres, "_time_PRSmixPlus.txt"), 
      row.names = F, sep = "\t", quote = F)
  }
}
writeLines("Finished")
return(0)
```


```{r}

dyslip_pgs_list <- "/bd-fs-mnt/TenantShare/exthome/home/l578311/prsmix_lists/inc_ischemic_PGS_list.txt"
combine_PRS(
    pheno_file = pheno_file ,
    covariate_file = covariate_file ,
    score_files_list = scores_file_vector ,
    trait_specific_score_file = dyslip_pgs_list ,
    pheno_name = pheno_name ,
    isbinary = TRUE , 
    out = "/bd-fs-mnt/TenantShare/exthome/home/l578311/pgs_sandbox/inc_ischemic_func_test_out", 
    liabilityR2 = TRUE  ,
    covar_list = covariates,
    cat_covar_list = categorical_vars,
    is_extract_adjSNPeff = FALSE,
    train_size_list = NULL,
    #power_thres_list = 0.95,
    pval_thres_list = (0.05/7),
    read_pred_training = FALSE,
    read_pred_testing = FALSE,
    original_beta_files_list = original_beta_files_list
  )
```
